---
title: "Data Cleaning"
author: "Grace Abels"
date: "4/19/2022"
output:
  pdf_document: default
  html_document: default
editor_options:
  chunk_output_type: console
---

```{r load packages, warning=FALSE, message=FALSE, echo = FALSE}
library(tidyverse)
library(knitr)
```

## Version 0 - All Unsorted Data

The version 0 dataset was pulled by Joel Luther. The dataset includes all entries from Claim Review that were published by sources based in the US (and in English). 14 unique fact-checking publishers were included. Included entries were those in which the reviewDate was between Jan 1st, 2016 and June 30th, 2021.

```{r v0 all unsorted, message=FALSE, warning=FALSE, echo = FALSE}

v0_all_us_unsorted <- read_csv("allUS_unsorted.csv")

v0_all_us_unsorted%>%
  count() %>%
  kable(caption = "All US Claims from ClaimReview 01/01/2016 - 06/30/2021")

v0_all_us_unsorted %>% 
  count(publisher.site)%>%
  kable(caption = "All US Claims from ClaimReview 01/01/2016 - 06/30/2021
        Sorted by Publisher")

```

### Version 1 - Only Political Figures

According to Asa Royal, we built a small data pipeline that tried to identify, for each claim, whether its uterrer was a person and, if they were, whether they met our definition of a politician. A politician, under our definition, was anyone who'd held or run for partisan office, or who'd been hired or appointed by a such a person to serve on a campaign or in a government agency.

We identified human claimants by feeding claimant names in our dataset through an entity recognition API. Given text like ""The Sierra Club"" or ""Bill Murray"", the API tried to detect what it referred to -- e.g. a person or an organization -- letting us label claimants as human.

Next, we tried to categorize human claimants as politicians or non-politicians. To do so, we fed claimants' names through a Wikipedia API. If a given claimant had a Wikipedia article about them, our code checked whether the article contained any ""politician indicators"". For example, if the infobox in an article about a claimant said they'd worked as a politician or political operative, we accepted that as fact and marked them as a politician. Likewise, if in the first few paragraphs of the Wikipedia article, text matched certain keywords (e.g. in a single sentence, the article mentioned its subject""ran"" in an ""election"" or ""work""ed in the""White House""), the code inferred that they, too, were a politician.

Additionally, once the code deduced that a claimant was a politician, it scanned their Wikipedia infobox for reference to their party affiliation, recording it in our dataset. If there was no Wikipedia article about a human claimant, as was occasionally the case for unsuccessful candidates for local office, the data pipeline tried to find a corresponding page for the claimant on Ballotpedia, an online encyclopedia of American politics. If it found one, it searched the page for the same markers looked for on Wikipedia pages, and similarly tried to identify the claimant's party affiliation. When the process above had been completed for each claimant, data collection was finished, and we filtered out all claims in the dataset whose claimants weren't labeled as human politicians. Following data collection, we undertook some light data cleaning. Namely, if we found that a claimant had been affiliated with several political parties, we gave them a party label corresponding to the party they'd identified with when they'd made the claim. In a small number of cases, our code categorized a claimant as a politician but failed to infer their party; those claimants' politician status and party ID were manually reviewed and edited at a later stage but were labelled as unknown_affiliation for the time being. The resulting dataset was of only claims made by politicians and political figures in the US with corresponding parties labeled.

```{r pull v1 data, message=FALSE, warning=FALSE, echo=FALSE}

v1_all_us_polclaims <- read_csv("all_us_pol_claims_updated_sep15.csv")

v1_all_us_polclaims %>%
  count() %>%
  kable(caption = "All Politician Claims (post-filter)")

v1_all_us_polclaims %>% 
  count(publisher.site) %>%
  kable(caption = "All Politician Claims (post-filter)
        Sorted by Publisher")

v1_all_us_polclaims %>% 
 count(claimant_party) %>%
 kable(caption = "All Politician Claims (post-filter)
        Sorted by Party")

```

### Version 2 - 5 Parties

At this stage we noticed that several of the claimants were assigned non-US political parties suggesting they were not US political figures. At this point, all claims whose party affiliation in claimant_party was not Democratic, Republican, Libertarian, Independent, or unknown_affilition were removed. Version 2 only consists of claims made by political figures belonging to U.S. political parties or ones that we assigned an unknown affiliation. At this time, those with unknown\_ affiliation were left unsorted. We addressed these once we had narrowed down to our smallest size datasets as to minimize hand sorting.

```{r create v2, message=FALSE, warning=FALSE, echo=FALSE}

v2_5parties <- v1_all_us_polclaims %>%
  filter(claimant_party == "Republican"|
           claimant_party == "Democratic"| 
         claimant_party == "Independent"|
         claimant_party == "Libertarian"| 
         claimant_party == "unknown_affiliation")

v2_5parties %>%
  count() %>%
  kable(caption = "All Claims in 5 Parties")

v2_5parties %>%
  count(publisher.site) %>%
  kable(caption = "All Claims in 5 Parties
        Sorted by Publisher")

v2_5parties %>%
  count(claimant_party) %>%
  kable(caption = "All Claims in 5 Parties
        Sorted by Party")

```

### Version 3 - Deduped

At this step we filtered out duplicate claims. These were claims in which both the url and and text of the claim were identical. The first appearance of each claim remained in the dataset, other were removed. This resulted in a database of 7,360 unique claims.\
\
*Note: Later on in our subject tagging we uncovered more duplicates that had identical text, but had been republished by a website under a different url. For transparency about our data cleaning process we are following our original steps and will remove those duplicates later.*

```{r dedupe for v3, message=FALSE, warning=FALSE, echo=FALSE}

v3_deduped <- v2_5parties %>%
  distinct(url, text, .keep_all = TRUE)

```

```{r display v3 results, message=FALSE, warning=FALSE, echo=FALSE}

v3_deduped %>%
  count() %>%
  kable(caption = "All Claims in 5 Parties - Deduped")

v3_deduped %>%
  count(publisher.site) %>%
  kable(caption = "All Claims in 5 Parties - Deduped and
        Sorted by Publisher")

v3_deduped %>%
  count(claimant_party) %>%
  kable(caption = "All Claims in 5 Parties - Deduped and
        Sorted by Party")

```

### Version 4 - Dejunked

For a final clean up before sorting by publisher and removing Trump, a member of our research team **manually identified 23 false positive rows** that either listed multiple claimants or had a false positive non-human claimant (e.g. "Donald Trump's campaign"). Those 23 claims were manually removed.

```{r manual removal for v4, message=FALSE, warning=FALSE, echo=FALSE}
v4_dejunked <- v3_deduped %>%
  filter(claimant != "Donald Trump For Prison") %>%
  filter(claimant != "Lauren Boebert; Rudy Giuliani") %>%
  filter(claimant != "Rick Scott's Starbuck's heckler") %>%
  filter(claimant != "Americans United for Change") %>%
  filter(claimant != "Consumers for Smart Solar") %>%
  filter(claimant != "Greg Gianforte's campaign") %>%
  filter(claimant != "Vietnam Veterans Against John McCain") %>%
  filter(claimant != "President Trump's lawyers") %>%
  filter(claimant != "Robin Vos and Scott Fitzgerald") %>%
  filter(claimant != "Robin Vos; Scott Fitzgerald") %>%
  filter(claimant != "Donald Trump 2020 Voters") %>%
  filter(claimant != "Michael Bloomberg; Joe Biden; Hillary Clinton; Adam Schiff") %>%
  filter(claimant != "Donald Trump and Mike Pence") %>%
  filter(claimant != "Bill DeBlasio and Brian Kemp") %>%
  filter(claimant != "John Roberts; Fox News correspondent") %>%
  filter(claimant != "The Trump campaign") %>%
  filter(claimant != "Sen. Ted Cruz (R-Texas) and Rep. Mark Meadows (R-N.C.)") %>%
  filter(claimant != "Keith Ellison spokesman") %>%
  filter(claimant != "Donald Trump ad") %>%
  filter(claimant != "Kamala Harris for the People") %>%
  filter(claimant != "Don Bolduc campaign") %>%
  filter(claimant != "Donald Trump campaign") %>%
  filter(claimant != "Joe Biden campaign") 

```

```{r tables for v4, message=FALSE, warning=FALSE, echo=FALSE}
v4_dejunked %>%
  count() %>%
  kable(caption = "All Claims in 5 Parties - Deduped and Dejunked")

v4_dejunked%>%
  count(publisher.site) %>%
  kable(caption = "All Claims in 5 Parties - Deduped and Dejunked
        Sorted by Publisher")

v4_dejunked %>%
  count(claimant_party) %>%
  kable(caption = "All Claims in 5 Parties - Deduped and Dejunked
        Sorted by Party")

```

At this stage, the dataset of 7,337 claims had been cleaned of irrelevant and redundant data removed. The dataset that remains consists of claims that were made by U.S. political figures, and fact-checked by one of 13 U.S based publishers.

### Version 5 - Split into Publisher data

At this stage, we began working with each publisher dataset separately. Only four publishers at this stage had enough claims to move forward: FactCheck.Org, The New York Times, PolitiFact, and The Washington Post. This resulted in four individual datasets, one per publisher.

```{r create v5, message=FALSE, warning=FALSE, echo=FALSE}
v5_pf <- v4_dejunked %>%
  filter(publisher.site == "politifact.com")

v5_wapo <- v4_dejunked %>%
  filter(publisher.site == "washingtonpost.com")

v5_fc <- v4_dejunked %>%
  filter(publisher.site == "factcheck.org")

v5_nyt <- v4_dejunked %>%
  filter(publisher.site == "nytimes.com")
```

```{r v5 tables, message=FALSE, warning=FALSE, echo=FALSE }
v5_pf %>%
  count() %>%
  kable(caption = "Politifact - All Claims in 5 Parties")

v5_pf %>%
  count(claimant_party) %>%
  kable(caption = "Politifact - All Claims in 5 Parties - 
        Sorted by Party")
v5_wapo %>%
  count() %>%
  kable(caption = "Washington Post - All Claims in 5 Parties")

v5_wapo %>%
  count(claimant_party) %>%
  kable(caption = "Washington Post - All Claims in 5 Parties - 
        Sorted by Party")

v5_fc %>%
  count() %>%
  kable(caption = "FactCheck.Org - All Claims in 5 Parties")

v5_fc %>%
  count(claimant_party) %>%
  kable(caption = "FactCheck.Org - All Claims in 5 Parties - 
        Sorted by Party")
v5_nyt %>%
  count() %>%
  kable(caption = "New York Times - All Claims in 5 Parties") 

v5_nyt %>%
  count(claimant_party) %>%
  kable(caption = "New York Times - All Claims in 5 Parties -  
        Sorted by Party")

```

### Version 6 - Removing Trump

We made the decision to remove Donald Trump entirely from this dataset. To do so, we first identified all the different iterations of his name within the `claimant` column. We could not simply remove all claimants whose names included Trump, because under our definitions, his daughter Ivanka Trump was a politician.

```{r find trump names, message=FALSE, warning=FALSE, echo=FALSE}

elim_trump <- v4_dejunked %>%
  filter(publisher.site == "politifact.com" |
           publisher.site ==  "washingtonpost.com" |
           publisher.site == "factcheck.org" |
           publisher.site == "nytimes.com")

elim_trump %>%
  filter(grepl('Trump|trump', claimant)) %>%
  group_by(claimant) %>%
  count() %>%
  kable(caption = "Names including 'Trump'")

```

With these four versions of Trump's name, we manually removed them from each V5 dataset. This created our V6 dataset without Trump for each publisher.

```{r create v6, message=FALSE, warning=FALSE, echo=FALSE}

v6_pf <- v5_pf %>%
  filter(claimant != "Donald J. Trump") %>%
  filter(claimant !=  "Donald trump") %>%
  filter(claimant != "Donald Trump") %>%
  filter(claimant != "President Donald J. Trump")

v6_wapo <- v5_wapo %>%
  filter(claimant != "Donald J. Trump") %>%
  filter(claimant !=  "Donald trump") %>%
  filter(claimant != "Donald Trump") %>%
  filter(claimant != "President Donald J. Trump")

v6_fc <- v5_fc %>%
  filter(claimant != "Donald J. Trump") %>%
  filter(claimant !=  "Donald trump") %>%
  filter(claimant != "Donald Trump") %>%
  filter(claimant != "President Donald J. Trump")

v6_nyt <- v5_nyt %>%
  filter(claimant != "Donald J. Trump") %>%
  filter(claimant !=  "Donald trump") %>%
  filter(claimant != "Donald Trump") %>%
  filter(claimant != "President Donald J. Trump")

```

```{r v6 tables, message=FALSE, warning=FALSE, echo=FALSE}
v6_pf %>%
  count() %>%
  kable(caption = "Politifact - Without Trump")

v6_pf %>%
  count(claimant_party) %>%
  kable(caption = "Politifact - Without Trump - 
        Sorted by Party")
v6_wapo %>%
  count() %>%
  kable(caption = "Washington Post - Without Trump")

v6_wapo %>%
  count(claimant_party) %>%
  kable(caption = "Washington Post - Without Trump - 
        Sorted by Party")

v6_fc %>%
  count() %>%
  kable(caption = "FactCheck.Org - Without Trump")

v6_fc %>%
  count(claimant_party) %>%
  kable(caption = "FactCheck.Org - Without Trump - 
        Sorted by Party")
v6_nyt %>%
  count() %>%
  kable(caption = "New York Times - Without Trump") 

v6_nyt %>%
  count(claimant_party) %>%
  kable(caption = "New York Times - Without Trump -  
        Sorted by Party")
```

### Version 7 - Sorting Unknown Affiliation

After we had the data divided into its smaller publisher sub-sets, and removed Trump from the data, our next step was to address anomalies in the claimant_party section. The code that was used to assign party affiliations scraped Wikipedia data and Ballotpedia for politicians and party affiliations. For about 100 different claims, the claimant was identified as a political figure but a party could not be found. They were labeled unknown_affiliation. We did not make an attempt to sort these previously, because there were simply too many. However, at this stage we have narrowed down to the data we plan to use. Therefore our next step was to manually assign party affiliations to the unknown_affiliation claims.

This was done by pulling the claims labeled unknown_affiliation for each publisher. Based on the name of the claimant and occasionally the url for the fact-check itself, claimants were either

1.  Assigned a party (Democrat, Republican, Independent, Libertarian)
2.  Marked to be removed from the dataset because they were not a political figure.

*NOTE: If a claimant served in an administration and spoke on behalf of that administration or politician, they were considered a political figure and assigned the affiliation of the administration or politician they served/spoke on behalf of. This applied to appointed cabinet members, political officials, and lawyers representing politicians. Political commentators, activists, and celebrities were removed and not considered political figures*

If you are reproducing this, these steps may not be relevant to you. Our filtering algorithm was unable to assign these political figures accurately despite many of them having political affiliations based on our definition. We are providing this information for the sake of transparency and to exemplify how one might manually sort the unknown_affiliations.

#### FactCheck.Org unknown affiliation review

First we reviewed the unknown_affiliation claims from the FactCheck.Org database.

```{r fc unknown_affilation table, message=FALSE, warning=FALSE,echo=FALSE}
v6_fc %>%
  filter(claimant_party == "unknown_affiliation") %>%
  group_by(claimant) %>% 
  count(sort = TRUE) %>%
  kable(caption = "Unknown Affiliation Claimants to be Sorted - FactCheck.Org")
```

There were 14 claims and 10 unique claimants.

After a manual review, the following claimants were assigned Republican affiliation:

Brett Giroir who [Served as Assistant Secretary of Health](https://en.wikipedia.org/wiki/Brett_Giroir) in the Trump Admin (2 claims)

Dr. Deborah Birx who [served in the Trump Admin](served%20in%20the%20Trump%20Admin)(1 claim)

John Kelly, [Former Chief of Staff in Trump Admin](https://en.wikipedia.org/wiki/John_F._Kelly) (1 claim)

Kevin K. McAleenan,  [Sec. of Homeland Sec. Trump Admin](https://en.wikipedia.org/wiki/Kevin_McAleenan) (1 claim)

Kirstjen Nielsen, [Homeland Security Sec. Trump Admin](https://en.wikipedia.org/wiki/Kirstjen_Nielsen) (4 claims)

Sidney Powell, [Lawyer representing President Donald Trump](https://en.wikipedia.org/wiki/Sidney_Powell) (1 claim)

Steve Cortes, [Former advisor to Donald Trump](https://premierespeakers.com/steve-cortes/bio) (1 claim) 

One claimant was assigned a Democratic affiliation.

Loretta Lynch, [Attorney General for Obama](https://en.wikipedia.org/wiki/Loretta_Lynch) (1 claim)

2 claimants were removed for not being political figures

Hal Turner, [commentator](https://en.wikipedia.org/wiki/Hal_Turner), Not a political figure (1 claim) 

Sean P. Conley, [Presidential physician](https://en.wikipedia.org/wiki/Sean_Conley), Not a political figure (1 claim) 

**In the end, 11 total claims were assigned a Republican affiliation, 1 claim was assigned a Democratic affiliation, and 2 claims were removed from the FactCheck.Org dataset.**

```{r v7 substitutions FC, message=FALSE, warning=FALSE, echo=FALSE}
v7_fc <- v6_fc %>%
  filter(claimant != "Hal Turner") %>%
  filter(claimant != "Sean P. Conley")
v7_fc$claimant_party[v7_fc$claimant == "Brett Giroir"] <- "Republican"
v7_fc$claimant_party[v7_fc$claimant == "Dr. Deborah Birx"] <- "Republican"
v7_fc$claimant_party[v7_fc$claimant == "John Kelly"] <- "Republican"
v7_fc$claimant_party[v7_fc$claimant == "Kevin K. McAleenan"] <- "Republican"
v7_fc$claimant_party[v7_fc$claimant == "Kirstjen Nielsen"] <- "Republican"
v7_fc$claimant_party[v7_fc$claimant == "Loretta Lynch"] <- "Democratic"
v7_fc$claimant_party[v7_fc$claimant == "Sidney Powell"] <- "Republican"
v7_fc$claimant_party[v7_fc$claimant == "Steve Cortes"] <- "Republican"

v7_fc %>%
  count(claimant_party) %>%
  kable(caption = "Factcheck.org by Claimant Party, 
        without Unknown Affiliation")
```

#### New York Times unknown affiliation review

Next we reviewed the unknown_affiliation claims for the New York Times database.

```{r}
v6_nyt %>%
  filter(claimant_party == "unknown_affiliation") %>%
  group_by(claimant) %>% 
  count(sort = TRUE) %>%
  kable(caption = "Unknown Affiliation Claimants to be Sorted - New York Times")
```

There were 2 unique claimants and 2 claims, both of which were assigned Republican Affiliation. 

Sidney Powell, [Lawyer representing President Donald Trump](https://en.wikipedia.org/wiki/Sidney_Powell) (1 claim)

Jesse Binnall, [Lawyer representing President Donald Trump](https://ballotpedia.org/Jesse_Binnall) (1 claim) \

```{r v7 substitutions NYT, message=FALSE, warning=FALSE, echo=FALSE}
v7_nyt <- v6_nyt
v7_nyt$claimant_party[v7_nyt$claimant == "Sidney Powell"] <- "Republican"
v7_nyt$claimant_party[v7_nyt$claimant == "Jesse Binnall"] <- "Republican"

v7_nyt %>%
  count(claimant_party) %>%
  kable(caption = "New York Times by Claimant Party, 
        without Unknown Affiliation")
```

#### PolitiFact unknown affiliation review

Next we reviewed unknown\_ affiliation claims for the PolitiFact database. There were 59 unique claimants and 72 claims.

```{r}
v6_pf %>%
  filter(claimant_party == "unknown_affiliation") %>%
  group_by(claimant) %>% 
  count(sort = TRUE) %>%
  kable(caption = "Unknown Affiliation Claimants to be Sorted - PolitiFact")
```

The following claimants were assigned a Republican affiliation:

Beth Parlato, [NY Republican Political candidate](https://ballotpedia.org/Beth_Parlato) (1 claim) 

Bob Spindell, [Republican member of WI elections commission](https://wisconsinexaminer.com/brief/spindell-reappointed-to-wisconsin-elections-commission/) (1 claim) 

Brett Giroir, [Served as Assistant Secretary of Health in the Trump Admin](https://en.wikipedia.org/wiki/Brett_Giroir) (2 claims) 

Brett McGurk, [National Security positions with Bush, trump, and Obama](https://en.wikipedia.org/wiki/Brett_McGurk) (1 claim) 

George Papadopoulos, [Served on trump campaign and Congressional candidate](https://en.wikipedia.org/wiki/George_Papadopoulos)(1 claim) 

H.R. McMaster, [National Security Advisor in Trump Admin](https://en.wikipedia.org/wiki/H._R._McMaster) (1 claim) 

Jesse Kremer, [Republican WI legislator](https://ballotpedia.org/Jesse_Kremer) (1 claim)

John Kelly,  [Former Chief of Staff in Trump Admin](https://en.wikipedia.org/wiki/John_F._Kelly) (3 claims)

Kirstjen Nielsen, [Homeland Security Sec. Trump Admin](https://en.wikipedia.org/wiki/Kirstjen_Nielsen) (1 claim)

Louis Marinelli, [California Political Candidate](https://en.wikipedia.org/wiki/Louis_J._Marinelli#The_President_of_California_Act) (1 claim) 

Marc Butler, [New York Republican](https://en.wikipedia.org/wiki/Marc_W._Butler) (1 claim) 

Paul Maner, [Georgia Republican Candidate](https://ballotpedia.org/Paul_Maner) (1 claim) 

Ryan Frazier, [Colorado Republican candidate](https://ballotpedia.org/Ryan_Frazier) (1 claim) 

Scott Dawson, [Alabama Republican Candidate](https://ballotpedia.org/Scott_Dawson_(Alabama)) (1 claim) 

Scott Jones, [Republican congressional candidate](https://www.politifact.com/factchecks/2016/oct/17/scott-jones/scott-jones-falsely-claims-judges-ruled-allegation/) (3 claims) 

Sidney Powell, [Lawyer representing President Donald Trump](https://en.wikipedia.org/wiki/Sidney_Powell) (2 claims)

Steve Cortes, [Former advisor to Donald Trump](https://premierespeakers.com/steve-cortes/bio) (1 claim) 

State Representatives, l[ink lists Wisconsin Assembly Republicans](https://www.politifact.com/factchecks/2019/mar/13/wisconsin-assembly-republicans/republicans-botch-claim-comparing-tuition-immigran/)) (1 claim) 

Todd Wilcox, [Florida Republican Candidate](https://ballotpedia.org/Todd_Wilcox_(Florida)) (1 claim)

The following claimants were assigned a Democratic affiliation:

Alejandro Mayorkas, [Biden Sec. of Homeland Security](https://www.dhs.gov/person/alejandro-mayorkas) (1 claim) 

Chris Meagher, [Dep. Press Sec. for Biden Admin](https://en.wikipedia.org/wiki/Chris_Meagher) (1 claim) 

Jeffery Zients, [Coronavirus Response Coordinator for Biden](https://en.wikipedia.org/wiki/Jeffrey_Zients) (1 claim) 

John Kirby, [Dept of defense Obama/Biden](https://en.wikipedia.org/wiki/John_Kirby_(admiral)), (1 claim) 

Loretta Lynch, [Attorney General for Obama](https://en.wikipedia.org/wiki/Loretta_Lynch) (1 claim) 

Mike Collier, [Dem. Texas Lieutenant Gov. candidate](https://ballotpedia.org/Mike_Collier_(Texas)) (1 claim) 

Nate McMurray, [NY Democratic politician](https://ballotpedia.org/Nate_McMurray) (1 claim) 

Clay Aiken, [NC Congressional Candidate](https://en.wikipedia.org/wiki/Clay_Aiken), (1 claim)\

The following claimants were assigned an Independent affiliation:

Alexander Strenger, [Ran for city council in Austin](https://ballotpedia.org/Alexander_Strenger) (1 claim) 

Bob Donovan, [non-partisan Milwaukee Common Council member](https://ballotpedia.org/Robert_G._Donovan_(Wisconsin)) (1 claim) 

Charles Francis, [Non-partisan Mayoral candidate](https://ballotpedia.org/Charles_Francis) (1 claim)

Lowell Holtz, [Non-partisan WI candidate](https://ballotpedia.org/Lowell_Holtz) (2 claims)

Michael Long, [New York State Conservative Party Chairman](https://www.politifact.com/personalities/michael-long/) (2 claims) 

William Taylor, [Ambassador to Ukraine](https://www.politifact.com/personalities/william-taylor/), worked in both party administration (1 claim) 

The following claimants were removed from the dataset:

Ashley Smith, [civilian](http://www.politifact.com/factchecks/2017/aug/03/ashley-smith/ashley-smith-transgender-residents-1-300-greg-abbo/), not political figure based on PF (1 claim)

Bill Maher, [political comedian](https://en.wikipedia.org/wiki/Bill_Maher), Not a political figure (1 claim) 

Bono, [singer](https://en.wikipedia.org/wiki/Bono), not political figure (1 claim)

Brit Hume, [journalist](https://en.wikipedia.org/wiki/Brit_Hume), not a political figure (1 claim) 

Charles Ramsey, [police chief](https://www.politifact.com/factchecks/2016/jul/10/charles-ramsey/how-many-police-departments-are-us/), not a political figure (1 claim) 

Daniel Kelly, [WI Supreme Court Justice](https://en.wikipedia.org/wiki/Daniel_Kelly_(Wisconsin_judge)), not a political figure (2 claims) 

David Martin,[ CEO of M CAM Inc., ](https://www.wipo.int/meetings/en/2006/scp_of_ge_06/speakers/martin.html)Not a political figure (1 claim) 

Edward Flynn, [Chief of Milwaukee Police dept.](https://en.wikipedia.org/wiki/Edward_A._Flynn), not a political figure (1 claim) 

Erin Burnett, [Political commentator,](https://en.wikipedia.org/wiki/Erin_Burnett) Not a political figure (1 claim)

Giffords, [Gun violence activist group](https://www.politifact.com/factchecks/2018/feb/22/giffords/florida-gov-rick-scott-made-it-illegal-doctors-tal/), not political figure (1 claim) 

Hal Turner, [commentator,](https://en.wikipedia.org/wiki/Hal_Turner) Not a political figure (1 claim) 

J. Christian Adams, [conservative activist](https://en.wikipedia.org/wiki/J._Christian_Adams), not a political figure (1 claim) 

Jake Tapper, [journalist](https://en.wikipedia.org/wiki/Jake_Tapper), Not a political figure (3 claims) 

James Hauser, [Not a political figure, only post](https://en.wikipedia.org/wiki/Jake_Tapper) (1 claim) 

John Patrick, [President of the Texas AFL-CIO](https://www.politifact.com/personalities/john-patrick/), not a political figure (1 claim) 

Kevin Downing, [Attorney to Paul Manafort](https://en.wikipedia.org/wiki/Kevin_M._Downing), not a political figure (1 claim) 

Lawrence O'Donnell,[ MSNBC Host](https://www.msnbc.com/the-last-word), not a political figure (1 claim) 

Lisa Moore, [English Professor](https://liberalarts.utexas.edu/english/faculty/moorell), not a political figure (1 claim) 

Maria Bartiromo, [FOX news host](https://en.wikipedia.org/wiki/Maria_Bartiromo), not a political figure (2 claims) 

Mark Levin,[ Talk show host](https://en.wikipedia.org/wiki/Mark_Levin), not a political figure (1 claim) 

Michael Screnock, [WI Supreme Court Justice](https://www.politifact.com/personalities/michael-screnock/), not a political figure (1 claim)

Mike Crute, [Radio show host](https://www.politifact.com/personalities/mike-crute/), not a political figure (1 claim)

Nicholas Burns, [Former ambassador, not a political figure](https://en.wikipedia.org/wiki/R._Nicholas_Burns) (2 claims) 

Ray Cross, [University President](https://ballotpedia.org/Ray_Cross), not a political figure (1 claim)

Robert Sanborn, [Nonprofit president ](https://childrenatrisk.org/staff/dr-robert-sanborn/), not a political figure (1 claim) 

Shelley Grogan, [WI Judge non-partisan](https://ballotpedia.org/Shelley_Grogan), not a political figure (1 claim) \

**25 Republican claims were assigned**

**8 Democratic claims were assigned** 

**8 Independent claims were assigned**

**31 claims were removed**

```{r v7 substitutions PF, message=FALSE, warning=FALSE, echo=FALSE}
v7_pf <- v6_pf %>%
  filter(claimant != "Ashley Smith") %>%
  filter(claimant != "Bill Maher") %>%
  filter(claimant != "Bono") %>%
  filter(claimant != "Brit Hume") %>%
  filter(claimant != "Charles Ramsey") %>%
  filter(claimant != "Daniel Kelly") %>%
  filter(claimant != "David Martin") %>%
  filter(claimant != "Edward Flynn") %>%
  filter(claimant != "Erin Burnett") %>%
  filter(claimant != "Giffords") %>%
  filter(claimant != "Hal Turner") %>%
  filter(claimant != "J. Christian Adams") %>%
  filter(claimant != "Jake Tapper") %>%
  filter(claimant != "James Hauser") %>%
  filter(claimant != "John Patrick") %>%
  filter(claimant != "Kevin Downing") %>%
  filter(claimant != "Lawrence O'Donnell") %>%
  filter(claimant != "Lisa Moore") %>%
  filter(claimant != "Maria Bartiromo") %>%
  filter(claimant != "Mark Levin") %>%
  filter(claimant != "Michael Screnock") %>%
  filter(claimant != "Mike Crute") %>%
  filter(claimant != "Nicholas Burns") %>%
  filter(claimant != "Ray Cross") %>%
  filter(claimant != "Robert Sanborn") %>%
  filter(claimant != "Shelley Grogan")
v7_pf$claimant_party[v7_pf$claimant == "Clay Aiken"] <- "Democratic"
v7_pf$claimant_party[v7_pf$claimant == "Alejandro Mayorkas"] <- "Democratic"
v7_pf$claimant_party[v7_pf$claimant == "Alexander Strenger"] <- "Independent"
v7_pf$claimant_party[v7_pf$claimant == "Beth Parlato"] <- "Republican"
v7_pf$claimant_party[v7_pf$claimant == "Bob Donovan"] <- "Independent"
v7_pf$claimant_party[v7_pf$claimant == "Bob Spindell"] <- "Republican"
v7_pf$claimant_party[v7_pf$claimant == "Brett Giroir"] <- "Republican"
v7_pf$claimant_party[v7_pf$claimant == "Brett McGurk"] <- "Republican"
v7_pf$claimant_party[v7_pf$claimant == "Charles Francis"] <- "Independent"
v7_pf$claimant_party[v7_pf$claimant == "Chris Meagher"] <- "Democratic"
v7_pf$claimant_party[v7_pf$claimant == "George Papadopoulos"] <- "Republican"
v7_pf$claimant_party[v7_pf$claimant == "H.R. McMaster"] <- "Republican"
v7_pf$claimant_party[v7_pf$claimant == "Jeffrey Zients"] <- "Democratic"
v7_pf$claimant_party[v7_pf$claimant == "Jesse Kremer"] <- "Republican"
v7_pf$claimant_party[v7_pf$claimant == "John Kelly"] <- "Republican"
v7_pf$claimant_party[v7_pf$claimant == "John Kirby"] <- "Democratic"
v7_pf$claimant_party[v7_pf$claimant == "Kirstjen Nielsen"] <- "Republican"
v7_pf$claimant_party[v7_pf$claimant == "Loretta Lynch"] <- "Democratic"
v7_pf$claimant_party[v7_pf$claimant == "Louis Marinelli"] <- "Republican"
v7_pf$claimant_party[v7_pf$claimant == "Lowell Holtz"] <- "Independent"
###THIS IS A DUPLICATE CLAIM
v7_pf$claimant_party[v7_pf$claimant == "Marc Butler"] <- "Republican"
v7_pf$claimant_party[v7_pf$claimant == "Michael Long"] <- "Independent"
v7_pf$claimant_party[v7_pf$claimant == "Mike Collier"] <- "Democratic"
v7_pf$claimant_party[v7_pf$claimant == "Nate McMurray"] <- "Democratic"
v7_pf$claimant_party[v7_pf$claimant == "Paul Maner"] <- "Republican"
v7_pf$claimant_party[v7_pf$claimant == "Ryan Frazier"] <- "Republican"
v7_pf$claimant_party[v7_pf$claimant == "Scott Dawson"] <- "Republican"
v7_pf$claimant_party[v7_pf$claimant == "Scott Jones"] <- "Republican"
v7_pf$claimant_party[v7_pf$claimant == "Sidney Powell"] <- "Republican"
v7_pf$claimant_party[v7_pf$claimant == "State representatives"] <- "Republican"
v7_pf$claimant_party[v7_pf$claimant == "Steve Cortes"] <- "Republican"
v7_pf$claimant_party[v7_pf$claimant == "Todd Wilcox"] <- "Republican"
v7_pf$claimant_party[v7_pf$claimant == "William Taylor"] <- "Independent"

v7_pf %>%
  count(claimant_party) %>%
  kable(caption = "PolitiFact by Claimant Party, 
        without Unknown Affiliation")
```

```{r v7 substitutions WP, message=FALSE, warning=FALSE, echo=FALSE}
v7_wapo <- v6_wapo %>%
  filter(claimant != "Anthony Fauci") %>%
  filter(claimant != "Jim Acosta")
v7_wapo$claimant_party[v7_wapo$claimant == "Charmaine Yoest"] <- "Republican"
v7_wapo$claimant_party[v7_wapo$claimant == "H.R. McMaster"] <- "Republican"
v7_wapo$claimant_party[v7_wapo$claimant == "Jim Mattis"] <- "Republican"
v7_wapo$claimant_party[v7_wapo$claimant == "Kirstjen Nielsen"] <- "Republican"
v7_wapo$claimant_party[v7_wapo$claimant == "Loretta Lynch"] <- "Democratic"
v7_wapo$claimant_party[v7_wapo$claimant == "Steve Cortes"] <- "Republican"
v7_wapo$claimant_party[v7_wapo$claimant == "Vanita Gupta"] <- "Democratic"

v7_wapo %>%
  count(claimant_party) %>%
  kable(caption = "Washington Post by Claimant Party, 
        without Unknown Affiliation")
```

### Version 8 - Removing Libertarian and Independents

```{r v8, message=FALSE, warning=FALSE, echo=FALSE}
v8_fc <- v7_fc %>%
  filter(claimant_party == "Democratic" |
           claimant_party == "Republican")

v8_fc %>%
  count(claimant_party) %>%
  kable(caption = "Factcheck.org by Claimant Party")

v8_nyt <- v7_nyt %>%
  filter(claimant_party == "Democratic" |
           claimant_party == "Republican")

v8_nyt %>%
  count(claimant_party) %>%
  kable(caption = "New York Times by Claimant Party")

v8_pf <- v7_pf %>%
  filter(claimant_party == "Democratic" |
           claimant_party == "Republican")

v8_pf %>%
  count(claimant_party) %>%
  kable(caption = "PolitiFact by Claimant Party")

v8_wapo <- v7_wapo %>%
  filter(claimant_party == "Democratic" |
           claimant_party == "Republican")

v8_wapo %>%
  count(claimant_party) %>%
  kable(caption = "Washington Post by Claimant Party")
```

```{r}
write_csv(v8_fc, file = "v8fc.csv")
write_csv(v8_nyt, file = "v8nyt.csv")
write_csv(v8_pf, file = "v8pf.csv")
write_csv(v8_wapo, file = "v8wapo.csv")
```
