---
title: "Data Cleaning"
output:
  pdf_document: default
  html_document: default
editor_options:
  chunk_output_type: console
  markdown: 
    wrap: sentence
---

```{r load packages, warning=FALSE, message=FALSE, echo = FALSE}
library(tidyverse)
library(knitr)
```

As described in the pages of the book \_\_\_\_\_\_\_\_\_\_, Bill Adair and his research team sought to understand how politicians lie through data.
Throughout our research, which is based on a dataset of existing fact-checks, we asked who lies more, who lies worse, what do they lie about, and where do they lie?

We originally theorized and experimented this process in spreadsheets, then translating our final process into R so that it may be reproduced and shared with the public.
We seek transparency in all our work, so feel free to explore the data yourself.
Sometimes the changes may seem small, irrelevant and tedious, but we believe it is vital to be fully transparent about each and every change we made.

This is the first of a series of Rmd files that walk a reader through our work.
In this document we describe our data management and cleaning process in detail.
We used a version system to keep ourselves organized and archive previous iterations of the data.

Each version represents a new dataset that has been altered from the previous one.

## Version 0 - All Unsorted Data

Our data was drawn from ClaimReview, a tagging system developed by the Duke Reporters Lab and Google in 2015.
When a fact-checker publishes a fact-check, they submit some baseline information to the ClaimReview database about the fact-check.
Some of the categories are listed below.

**`url`** - Link to the fact-check.

**`title`** - The title of the fact-check, often a summary.

**`publisher.name`** - Name of the site that published the fact-check (PolitiFact).

**`reviewDate`** - Date the fact-check was published.

**`text`** - The actual claim being fact-checked.

**`claimDate`** - The date the claim was made.

**`claimant`** - The name of the speaker/person making the claim.

**`textualRating`** - The rating assigned to the claim.

ClaimReview was initially developed to promote fact-checks in search results, incentivizing fact-checkers to participate.
ClaimReview also had something of an unintended benefit: It created a giant, growing database of fact-checks uploaded by fact-checkers around the world.
From US-based fact-checkers alone, there are 18,770 claims from January 2016 to June 2021, the timeline in question for our analysis.

The version 0 dataset was pulled by Joel Luther in the Duke Reporters Lab.
The dataset includes all entries from ClaimReview that were published by fact checkers based in the US (and in English).
14 unique fact-checking publishers were included.
Included entries were those in which the reviewDate was between Jan 1st, 2016 and June 30th, 2021.

```{r v0 all unsorted, message=FALSE, warning=FALSE, echo = FALSE}

v0_all_us_unsorted <- read_csv("allUS_unsorted.csv")

v0_all_us_unsorted%>%
  count() %>%
  kable(caption = "V0, Original ClaimReview Data")

v0_all_us_unsorted %>% 
  count(publisher.site)%>%
  kable(caption = "V0, Sorted by Publisher")

```

To access V0, and any other code or datasets used in this project, feel free to visit our [public github repository](https://github.com/gracelabels/PatternsofLying).

### Version 1 - Only Political Figures

ClaimReview contains fact-checks for all sorts of misinformation, not just the lies of politicians, which was the focus of our research.
So we narrowed down the data by claimant to just claims made by political figures.
That process is described below.

Asa Royal, the software engineer assisting on the project, built a data pipeline that tried to identify whether the speaker of each claim was a person and whether they met our definition of a politician.
**For our purposes, a politician was anyone who'd held or run for partisan office, or who'd been hired or appointed by a such a person to serve on a campaign or in a government agency. For the remainder of our analysis we will be using the terms politician and political figure interchangably.**

We identified human claimants by feeding claimant names in our dataset through an entity recognition API.
Given text like ""The Sierra Club"" or ""Bill Murray"", the API tried to detect what it referred to -- *e.g.* a person or an organization -- letting us label claimants as human.

Next, we categorized human claimants as politicians or non-politicians.
We began by running claimant names through a Wikipedia API. If a given claimant had a Wikipedia article about them, our code checked whether the article contained any "politician indicators." For example, if the infobox in an article about a claimant said they'd worked as a politician or political operative, we accepted that as fact and marked them as a politician.
Likewise, if in the first few paragraphs of the Wikipedia article, text matched certain keywords (*e.g.* in a single sentence, the article mentioned its subject "ran" in an "election" or "work"ed in the "White House"), the code inferred that they, too, were a politician.

Additionally, once the code deduced that a claimant was a politician, it scanned their Wikipedia infobox for reference to their party affiliation, recording it in our dataset under `claimant_party`.
If there was no Wikipedia article about a human claimant, as was occasionally the case for unsuccessful candidates for local office, the data pipeline tried to find a corresponding page for the claimant on [Ballotpedia](https://ballotpedia.org/Main_Page), an online encyclopedia of American politics.
If it found one, it searched the page for the same markers looked for on Wikipedia pages, and similarly tried to identify the claimant's party affiliation.

Once the aforementioned process was complete for each claimant and we had filtered out all claims in the dataset whose claimants weren't labeled as human politicians, we considered data collection to be finished.
To begin making this data workable for our research goals, we undertook some light data cleaning.
Namely, if we found that a claimant had been affiliated with several political parties, we gave them a party label corresponding to the party they were associated with when they'd made the claim.
In a small number of cases, our code categorized a claimant as a politician but failed to infer their party; those claimants' politician status and party ID were manually reviewed and edited at a later stage but were labelled as unknown_affiliation for the time being.
The resulting dataset only contained claims by political figures with a corresponding party label.

```{r pull v1 data, message=FALSE, warning=FALSE, echo=FALSE}

v1_all_us_polclaims <- read_csv("all_US_politician_claims.csv")

v1_all_us_polclaims %>%
  count() %>%
  kable(caption = "V1, All Politician Claims")

v1_all_us_polclaims %>% 
  count(publisher.site) %>%
  kable(caption = "V1, Sorted by Publisher")

v1_all_us_polclaims %>%
  count(claimant_party) %>%
  arrange() %>% 
  kable(caption = "V1, Sorted by Party")

```

### Version 2 - 5 Parties

As seen in Table 5, several of the claimants were assigned non-US political parties, suggesting they were not US political figures.
At this point, all claims whose party affiliation in `claimant_party` was not Democratic, Republican, Libertarian, Independent, or unknown_affilition were removed.
Version 2 only consists of claims made by political figures belonging to U.S. political parties or ones that we assigned an unknown affiliation.
At this time, those with unknown\_ affiliation were left unsorted.
With further cleaning needing to be done, we held off on manually sorting these claims until we had further narrowed our dataset.

```{r create v2, message=FALSE, warning=FALSE, echo=FALSE}

v2_5parties <- v1_all_us_polclaims %>%
  filter(claimant_party == "Republican"|
           claimant_party == "Democratic"| 
         claimant_party == "Independent"|
         claimant_party == "Libertarian"| 
         claimant_party == "unknown_affiliation")

v2_5parties %>%
  count() %>%
  kable(caption = "V2, All Claimants in 5 Parties")

v2_5parties %>%
  count(publisher.site) %>%
  kable(caption = "V2, Sorted by Publisher")

v2_5parties %>%
  count(claimant_party) %>%
  kable(caption = "V2, Sorted by Party")

```

### Version 3 - Deduped

We also wanted to check for duplicate claims.
To do so, we filtered out claims where both the `url` and `text` of the claim were identical.
The first appearance of each claim remained in the dataset, others were removed.
This resulted in a database of 7,360 unique claims.\
\
*Note: Later on in our subject tagging we discovered more duplicates that had identical text, but had been republished by a website under a different url, so they were not deleted in this step. For transparency about our data cleaning process we are following our original steps and will remove those duplicates later.*

```{r dedupe for v3, message=FALSE, warning=FALSE, echo=FALSE}

v3_deduped <- v2_5parties %>%
  distinct(url, text, .keep_all = TRUE)

```

```{r display v3 results, message=FALSE, warning=FALSE, echo=FALSE}

v3_deduped %>%
  count() %>%
  kable(caption = "V3, Duplicates Removed")

v3_deduped %>%
  count(publisher.site) %>%
  kable(caption = "V3, Sorted by Publisher")

v3_deduped %>%
  count(claimant_party) %>%
  kable(caption = "V3, Sorted by Party")

```

### Version 4 - Dejunked

In a final clean up before sorting by publisher and removing Trump, a member of our research team **manually identified 23 false positive rows** that either listed multiple claimants or had a false positive non-human claimant (e.g. "Donald Trump's campaign").
Those 23 claims were manually removed.

```{r manual removal for v4, message=FALSE, warning=FALSE, echo=FALSE}
v4_dejunked <- v3_deduped %>%
  filter(claimant != "Donald Trump For Prison") %>%
  filter(claimant != "Lauren Boebert; Rudy Giuliani") %>%
  filter(claimant != "Rick Scott's Starbuck's heckler") %>%
  filter(claimant != "Americans United for Change") %>%
  filter(claimant != "Consumers for Smart Solar") %>%
  filter(claimant != "Greg Gianforte's campaign") %>%
  filter(claimant != "Vietnam Veterans Against John McCain") %>%
  filter(claimant != "President Trump's lawyers") %>%
  filter(claimant != "Robin Vos and Scott Fitzgerald") %>%
  filter(claimant != "Robin Vos; Scott Fitzgerald") %>%
  filter(claimant != "Donald Trump 2020 Voters") %>%
  filter(claimant != "Michael Bloomberg; Joe Biden; Hillary Clinton; Adam Schiff") %>%
  filter(claimant != "Donald Trump and Mike Pence") %>%
  filter(claimant != "Bill DeBlasio and Brian Kemp") %>%
  filter(claimant != "John Roberts; Fox News correspondent") %>%
  filter(claimant != "The Trump campaign") %>%
  filter(claimant != "Sen. Ted Cruz (R-Texas) and Rep. Mark Meadows (R-N.C.)") %>%
  filter(claimant != "Keith Ellison spokesman") %>%
  filter(claimant != "Donald Trump ad") %>%
  filter(claimant != "Kamala Harris for the People") %>%
  filter(claimant != "Don Bolduc campaign") %>%
  filter(claimant != "Donald Trump campaign") %>%
  filter(claimant != "Joe Biden campaign") 

```

```{r tables for v4, message=FALSE, warning=FALSE, echo=FALSE}
v4_dejunked %>%
  count() %>%
  kable(caption = "V4, Removal of False Positives")

v4_dejunked%>%
  count(publisher.site) %>%
  kable(caption = "V4, Sorted by Publisher")

v4_dejunked %>%
  count(claimant_party) %>%
  kable(caption = "V4, Sorted by Party")
```

At this stage, the dataset of 7,337 claims had been filtered to fit the parameters of our query.
The dataset that remains consists of claims that were made by U.S. political figures, and fact-checked by one of 13 U.S based publishers.

### Version 5 - Split into Publisher data

From here on out, we began working with each publisher dataset separately to retain the integrity of the different rating systems.
Each publisher uses a unique scale for its ratings and has its own methodology for assigning ratings.
Therefore it would not have been appropriate to equate a PolitiFact "mostly false" with a WashingtonPost "Three Pinocchios" rating.
As such, we had to analyze each publisher individually.
Only four publishers had enough remaining claims to move forward: FactCheck.Org, The New York Times, PolitiFact, and The Washington Post.
We divided the data into four individual datasets, one per publisher.

```{r create v5, message=FALSE, warning=FALSE, echo=FALSE}
v5_fc <- v4_dejunked %>%
  filter(publisher.site == "factcheck.org")

v5_nyt <- v4_dejunked %>%
  filter(publisher.site == "nytimes.com")

v5_pf <- v4_dejunked %>%
  filter(publisher.site == "politifact.com")

v5_wapo <- v4_dejunked %>%
  filter(publisher.site == "washingtonpost.com")
```

```{r v5 tables, message=FALSE, warning=FALSE, echo=FALSE }
v5_fc %>%
  count() %>%
  kable(caption = "V5, FactCheck.Org")

v5_fc %>%
  count(claimant_party) %>%
  kable(caption = "V5, FactCheck.Org, Sorted by Party")

v5_nyt %>%
  count() %>%
  kable(caption = "V5, New York Times") 

v5_nyt %>%
  count(claimant_party) %>%
  kable(caption = "V5, New York Times, Sorted by Party")

v5_pf %>%
  count() %>%
  kable(caption = "V5, PolitiFact")

v5_pf %>%
  count(claimant_party) %>%
  kable(caption = "V5, PolitiFact, Sorted by Party")

v5_wapo %>%
  count() %>%
  kable(caption = "V5, Washington Post")

v5_wapo %>%
  count(claimant_party) %>%
  kable(caption = "V5, Washington Post, Sorted by Party")
```

### Version 6 - Removing Trump

We made the decision to remove Donald Trump entirely from all four datasets.
In V4, before we split the data by publisher, Trump's statements accounted for 2984 of the 7337 total claims in the data, a significant proportion of which were falsehoods.
Responsible for 40% of all claims fact-checked, he was overrepresented in the data.
Accounting for over 1/3 of the dataset, all of Trump's claims were attributed to Republicans when we compared the two parties.
The vastly disproportionate number of Republican falsehoods threatened to skew party analyses, solely due to his presence in the data.
To do so, we first identified all the different iterations of "Trump" within the `claimant` column.
We could not simply remove all claimants whose names included Trump, because by our definitions, his daughter Ivanka Trump was a political figure.

```{r find trump names, message=FALSE, warning=FALSE, echo=FALSE}

elim_trump <- v4_dejunked %>%
  filter(publisher.site == "politifact.com" |
           publisher.site ==  "washingtonpost.com" |
           publisher.site == "factcheck.org" |
           publisher.site == "nytimes.com")

elim_trump %>%
  filter(grepl('Trump|trump', claimant)) %>%
  group_by(claimant) %>%
  count() %>%
  kable(caption = "Names including 'Trump'")

```

With these four versions of Donald Trump's name, we manually removed them from the data.
This created a dataset without Trump for each publisher, denoted V6.

```{r create v6, message=FALSE, warning=FALSE, echo=FALSE}

v6_pf <- v5_pf %>%
  filter(claimant != "Donald J. Trump") %>%
  filter(claimant !=  "Donald trump") %>%
  filter(claimant != "Donald Trump") %>%
  filter(claimant != "President Donald J. Trump")

v6_wapo <- v5_wapo %>%
  filter(claimant != "Donald J. Trump") %>%
  filter(claimant !=  "Donald trump") %>%
  filter(claimant != "Donald Trump") %>%
  filter(claimant != "President Donald J. Trump")

v6_fc <- v5_fc %>%
  filter(claimant != "Donald J. Trump") %>%
  filter(claimant !=  "Donald trump") %>%
  filter(claimant != "Donald Trump") %>%
  filter(claimant != "President Donald J. Trump")

v6_nyt <- v5_nyt %>%
  filter(claimant != "Donald J. Trump") %>%
  filter(claimant !=  "Donald trump") %>%
  filter(claimant != "Donald Trump") %>%
  filter(claimant != "President Donald J. Trump")

```

```{r v6 tables, message=FALSE, warning=FALSE, echo=FALSE}
v6_fc %>%
  count() %>%
  kable(caption = "V6, FactCheck.Org without D. Trump")

v6_fc %>%
  count(claimant_party) %>%
  kable(caption = "V6, FactCheck.Org, Sorted by Party")
v6_nyt %>%
  count() %>%
  kable(caption = "V6, New York Times without D. Trump") 

v6_nyt %>%
  count(claimant_party) %>%
  kable(caption = "V6, New York Times, Sorted by Party")

v6_pf %>%
  count() %>%
  kable(caption = "V6, PolitiFact without D. Trump")

v6_pf %>%
  count(claimant_party) %>%
  kable(caption = "V6, PolitiFact, Sorted by Party")
v6_wapo %>%
  count() %>%
  kable(caption = "V6, Washington Post without D. Trump")

v6_wapo %>%
  count(claimant_party) %>%
  kable(caption = "V6, Washington Post, Sorted by Party")
```

From this point forward, Donald Trump's claims are not included in any analysis of the data.

### Version 7 - Sorting Unknown Affiliation

After we had divided the data into smaller publisher sub-sets and removed Trump from the data, our next step was to address missing data in the `claimant_party` column.
As a reminder, the code that was used to assign party affiliations scraped Wikipedia data and Ballotpedia for politicians and party affiliations.
Across the four datasets, 100 claims remained that were made by a claimant who was identified as a political figure but for whom a party could not be found.
They were labeled unknown_affiliation.
Having reduced our data to the selection of claims we planned to analyze, our next step was to manually assign these claimants a party affiliations.

This was done by pulling the claims labeled unknown_affiliation for each publisher.
Based on the name of the claimant and the content of the fact-check, claimants were either

1.  Assigned a party (Democrat, Republican, Independent, Libertarian)
2.  Removed from the dataset because they were not a political figure.

*NOTE: If a claimant served in an administration and spoke on behalf of that administration or politician, they were considered a political figure and assigned the affiliation of the administration or politician they served/spoke on behalf of. This may not always align with their personal political beliefs but represented the party they spoke for within their role. This applied to appointed cabinet members, political officials, and lawyers representing politicians. Political commentators, activists, and celebrities were removed and not considered political figures*

If you are reproducing this, these steps may not be relevant to you.
Our filtering algorithm was unable to assign these political figures accurately despite many of them having political affiliations based on our definition.
We are providing this information for the sake of transparency and to exemplify how one might manually sort the unknown_affiliations.

#### FactCheck.Org unknown affiliation review

```{r fc unknown_affilation table, message=FALSE, warning=FALSE,echo=FALSE}
v6_fc %>%
  filter(claimant_party == "unknown_affiliation") %>%
  group_by(claimant) %>% 
  count(sort = TRUE) %>%
  kable(caption = "V6, FactCheck.Org including unknown affiliation")
```

There were 14 claims and 10 unique claimants.

**After a manual review, the** **following claimants were assigned Republican affiliation:**

Brett Giroir who [Served as Assistant Secretary of Health](https://en.wikipedia.org/wiki/Brett_Giroir) in the Trump Admin (2 claims)

Dr. Deborah Birx who [served in the Trump admin](served%20in%20the%20Trump%20Admin)(1 claim)

John Kelly, [Former Chief of Staff in Trump admin](https://en.wikipedia.org/wiki/John_F._Kelly) (1 claim)

Kevin K. McAleenan, [Sec. of Homeland Sec. Trump admin](https://en.wikipedia.org/wiki/Kevin_McAleenan) (1 claim)

Kirstjen Nielsen, [Homeland Security Sec. Trump admin](https://en.wikipedia.org/wiki/Kirstjen_Nielsen) (4 claims)

Sidney Powell, [Lawyer representing President Donald Trump](https://en.wikipedia.org/wiki/Sidney_Powell) (1 claim)

Steve Cortes, [Former advisor to Donald Trump](https://premierespeakers.com/steve-cortes/bio) (1 claim)

**One claimant was assigned a Democratic affiliation:**

Loretta Lynch, [Attorney General for Obama](https://en.wikipedia.org/wiki/Loretta_Lynch) (1 claim)

**2 claimants were removed for not being political figures:**

Hal Turner, [commentator](https://en.wikipedia.org/wiki/Hal_Turner), not a political figure (1 claim)

Sean P. Conley, [Presidential physician](https://en.wikipedia.org/wiki/Sean_Conley), not a political figure (1 claim)

**In the end, 11 total claims were assigned a Republican affiliation, 1 claim was assigned a Democratic affiliation, and 2 claims were removed from the FactCheck.Org dataset.** Table 33 shows the count by party after this adjustment.

```{r v7 substitutions FC, message=FALSE, warning=FALSE, echo=FALSE}
v7_fc <- v6_fc %>%
  filter(claimant != "Hal Turner") %>%
  filter(claimant != "Sean P. Conley")
v7_fc$claimant_party[v7_fc$claimant == "Brett Giroir"] <- "Republican"
v7_fc$claimant_party[v7_fc$claimant == "Dr. Deborah Birx"] <- "Republican"
v7_fc$claimant_party[v7_fc$claimant == "John Kelly"] <- "Republican"
v7_fc$claimant_party[v7_fc$claimant == "Kevin K. McAleenan"] <- "Republican"
v7_fc$claimant_party[v7_fc$claimant == "Kirstjen Nielsen"] <- "Republican"
v7_fc$claimant_party[v7_fc$claimant == "Loretta Lynch"] <- "Democratic"
v7_fc$claimant_party[v7_fc$claimant == "Sidney Powell"] <- "Republican"
v7_fc$claimant_party[v7_fc$claimant == "Steve Cortes"] <- "Republican"

v7_fc %>%
  count(claimant_party) %>%
  kable(caption = "V7, FactCheck.Org, corrected party affiliation")
```

#### New York Times unknown affiliation review

```{r nyt unknown affiliation table, message=FALSE, warning=FALSE, echo=FALSE}
v6_nyt %>%
  filter(claimant_party == "unknown_affiliation") %>%
  group_by(claimant) %>% 
  count(sort = TRUE) %>%
  kable(caption = "V6, New York Times, including unknown affiliation")
```

**There were 2 unique claimants and 2 claims, both of which were assigned Republican affiliation:**

Sidney Powell, [Lawyer representing President Donald Trump](https://en.wikipedia.org/wiki/Sidney_Powell) (1 claim)

Jesse Binnall, [Lawyer representing President Donald Trump](https://ballotpedia.org/Jesse_Binnall) (1 claim)\

**2 total claims were assigned a Republican affiliation** Table 35 shows the count by party after this adjustment.

```{r v7 substitutions NYT, message=FALSE, warning=FALSE, echo=FALSE}
v7_nyt <- v6_nyt
v7_nyt$claimant_party[v7_nyt$claimant == "Sidney Powell"] <- "Republican"
v7_nyt$claimant_party[v7_nyt$claimant == "Jesse Binnall"] <- "Republican"

v7_nyt %>%
  count(claimant_party) %>%
  kable(caption = "V7, New York Times, corrected party affiliation")
```

#### PolitiFact unknown affiliation review

```{r pf unknown affiliation table, message=FALSE, warning=FALSE, echo=FALSE }
v6_pf %>%
  filter(claimant_party == "unknown_affiliation") %>%
  group_by(claimant) %>% 
  count(sort = TRUE) %>%
  kable(caption = "V6, PolitiFact, including unknown affiliation")
```

There were 59 unique claimants and 72 claims.

**The following claimants were assigned a Republican affiliation:**

Beth Parlato, [NY Republican political candidate](https://ballotpedia.org/Beth_Parlato) (1 claim)

Bob Spindell, [Republican member of WI elections commission](https://wisconsinexaminer.com/brief/spindell-reappointed-to-wisconsin-elections-commission/) (1 claim)

Brett Giroir, [Served as Assistant Secretary of Health in the Trump admin](https://en.wikipedia.org/wiki/Brett_Giroir) (2 claims)

Brett McGurk, [National Security positions with Bush, Trump, and Obama](https://en.wikipedia.org/wiki/Brett_McGurk) (1 claim)

George Papadopoulos, [Served on Trump campaign and congressional candidate](https://en.wikipedia.org/wiki/George_Papadopoulos)(1 claim)

H.R. McMaster, [National Security Advisor in Trump admin](https://en.wikipedia.org/wiki/H._R._McMaster) (1 claim)

Jesse Kremer, [Republican WI legislator](https://ballotpedia.org/Jesse_Kremer) (1 claim)

John Kelly, [Former Chief of Staff in Trump Admin](https://en.wikipedia.org/wiki/John_F._Kelly) (3 claims)

Kirstjen Nielsen, [Homeland Security Sec. Trump Admin](https://en.wikipedia.org/wiki/Kirstjen_Nielsen) (1 claim)

Louis Marinelli, [California Political Candidate](https://en.wikipedia.org/wiki/Louis_J._Marinelli#The_President_of_California_Act) (1 claim)

Marc Butler, [New York Republican](https://en.wikipedia.org/wiki/Marc_W._Butler) (1 claim)

Paul Maner, [Georgia Republican candidate](https://ballotpedia.org/Paul_Maner) (1 claim)

Ryan Frazier, [Colorado Republican candidate](https://ballotpedia.org/Ryan_Frazier) (1 claim)

Scott Dawson, [Alabama Republican candidate](https://ballotpedia.org/Scott_Dawson_(Alabama)) (1 claim)

Scott Jones, [Republican congressional candidate](https://www.politifact.com/factchecks/2016/oct/17/scott-jones/scott-jones-falsely-claims-judges-ruled-allegation/) (3 claims)

Sidney Powell, [Lawyer representing President Donald Trump](https://en.wikipedia.org/wiki/Sidney_Powell) (2 claims)

Steve Cortes, [Former advisor to Donald Trump](https://premierespeakers.com/steve-cortes/bio) (1 claim) 

State Representatives, l[ink lists Wisconsin Assembly Republicans](https://www.politifact.com/factchecks/2019/mar/13/wisconsin-assembly-republicans/republicans-botch-claim-comparing-tuition-immigran/) (1 claim)

Todd Wilcox, [Florida Republican candidate](https://ballotpedia.org/Todd_Wilcox_(Florida)) (1 claim)

**The following claimants were assigned a Democratic affiliation:**

Alejandro Mayorkas, [Biden Sec. of Homeland Security](https://www.dhs.gov/person/alejandro-mayorkas) (1 claim)

Chris Meagher, [Dep. Press Sec. for Biden Admin](https://en.wikipedia.org/wiki/Chris_Meagher) (1 claim)

Jeffery Zients, [Coronavirus Response Coordinator for Biden](https://en.wikipedia.org/wiki/Jeffrey_Zients) (1 claim)

John Kirby, [Dept of Defense Obama/Biden](https://en.wikipedia.org/wiki/John_Kirby_(admiral)), (1 claim)

Loretta Lynch, [Attorney General for Obama](https://en.wikipedia.org/wiki/Loretta_Lynch) (1 claim)

Mike Collier, [Dem. Texas Lieutenant Gov. candidate](https://ballotpedia.org/Mike_Collier_(Texas)) (1 claim)

Nate McMurray, [NY Democratic politician](https://ballotpedia.org/Nate_McMurray) (1 claim)

Clay Aiken, [NC Congressional candidate](https://en.wikipedia.org/wiki/Clay_Aiken), (1 claim)\

**The following claimants were assigned an Independent affiliation:**

Alexander Strenger, [ran for city council in Austin](https://ballotpedia.org/Alexander_Strenger) (1 claim)

Bob Donovan, [non-partisan Milwaukee Common Council member](https://ballotpedia.org/Robert_G._Donovan_(Wisconsin)) (1 claim)

Charles Francis, [non-partisan Mayoral candidate](https://ballotpedia.org/Charles_Francis) (1 claim)

Lowell Holtz, [non-partisan WI candidate](https://ballotpedia.org/Lowell_Holtz) (2 claims)

Michael Long, [New York State Conservative Party Chairman](https://www.politifact.com/personalities/michael-long/) (2 claims)

William Taylor, [Ambassador to Ukraine](https://www.politifact.com/personalities/william-taylor/), worked in both party administrations (1 claim)

**The following claimants were removed from the dataset:**

Ashley Smith, [civilian](http://www.politifact.com/factchecks/2017/aug/03/ashley-smith/ashley-smith-transgender-residents-1-300-greg-abbo/), not political figure based on PF (1 claim)

Bill Maher, [political comedian](https://en.wikipedia.org/wiki/Bill_Maher), not a political figure (1 claim)

Bono, [singer](https://en.wikipedia.org/wiki/Bono), not political figure (1 claim)

Brit Hume, [journalist](https://en.wikipedia.org/wiki/Brit_Hume), not a political figure (1 claim)

Charles Ramsey, [police chief](https://www.politifact.com/factchecks/2016/jul/10/charles-ramsey/how-many-police-departments-are-us/), not a political figure (1 claim)

Daniel Kelly, [WI Supreme Court Justice](https://en.wikipedia.org/wiki/Daniel_Kelly_(Wisconsin_judge)), not a political figure (2 claims)

David Martin, [CEO of M CAM Inc.,](https://www.wipo.int/meetings/en/2006/scp_of_ge_06/speakers/martin.html) not a political figure (1 claim)

Edward Flynn, [Chief of Milwaukee Police dept.](https://en.wikipedia.org/wiki/Edward_A._Flynn), not a political figure (1 claim)

Erin Burnett, [Political commentator,](https://en.wikipedia.org/wiki/Erin_Burnett) not a political figure (1 claim)

Giffords, [Gun violence activist group](https://www.politifact.com/factchecks/2018/feb/22/giffords/florida-gov-rick-scott-made-it-illegal-doctors-tal/), not political figure (1 claim)

Hal Turner, [commentator,](https://en.wikipedia.org/wiki/Hal_Turner) not a political figure (1 claim)

J. Christian Adams, [conservative activist](https://en.wikipedia.org/wiki/J._Christian_Adams), not a political figure (1 claim)

Jake Tapper, [journalist](https://en.wikipedia.org/wiki/Jake_Tapper), not a political figure (3 claims)

James Hauser, [civilian, not a political figure, only post](https://en.wikipedia.org/wiki/Jake_Tapper) (1 claim)

John Patrick, [President of the Texas AFL-CIO](https://www.politifact.com/personalities/john-patrick/), not a political figure (1 claim)

Kevin Downing, [Attorney to Paul Manafort](https://en.wikipedia.org/wiki/Kevin_M._Downing), not a political figure (1 claim)

Lawrence O'Donnell, [MSNBC Host](https://www.msnbc.com/the-last-word), not a political figure (1 claim)

Lisa Moore, [English Professor](https://liberalarts.utexas.edu/english/faculty/moorell), not a political figure (1 claim)

Maria Bartiromo, [FOX news host](https://en.wikipedia.org/wiki/Maria_Bartiromo), not a political figure (2 claims)

Mark Levin, [Talk show host](https://en.wikipedia.org/wiki/Mark_Levin), not a political figure (1 claim)

Michael Screnock, [WI Supreme Court Justice](https://www.politifact.com/personalities/michael-screnock/), not a political figure (1 claim)

Mike Crute, [radio show host](https://www.politifact.com/personalities/mike-crute/), not a political figure (1 claim)

Nicholas Burns, [Former ambassador, not a political figure](https://en.wikipedia.org/wiki/R._Nicholas_Burns) (2 claims)

Ray Cross, [University President](https://ballotpedia.org/Ray_Cross), not a political figure (1 claim)

Robert Sanborn, [nonprofit president](https://childrenatrisk.org/staff/dr-robert-sanborn/), not a political figure (1 claim)

Shelley Grogan, [WI judge non-partisan](https://ballotpedia.org/Shelley_Grogan), not a political figure (1 claim)\

**In the end, 25 claims were assigned a Republican affiliation, 8 claims were assigned a Democratic affiliation, 8 claims were assigned an Independent affiliation, and 31 claims were removed from the PolitiFact dataset.** Table 37 shows the counts by party after the sort.

```{r v7 substitutions PF, message=FALSE, warning=FALSE, echo=FALSE}
v7_pf <- v6_pf %>%
  filter(claimant != "Ashley Smith") %>%
  filter(claimant != "Bill Maher") %>%
  filter(claimant != "Bono") %>%
  filter(claimant != "Brit Hume") %>%
  filter(claimant != "Charles Ramsey") %>%
  filter(claimant != "Daniel Kelly") %>%
  filter(claimant != "David Martin") %>%
  filter(claimant != "Edward Flynn") %>%
  filter(claimant != "Erin Burnett") %>%
  filter(claimant != "Giffords") %>%
  filter(claimant != "Hal Turner") %>%
  filter(claimant != "J. Christian Adams") %>%
  filter(claimant != "Jake Tapper") %>%
  filter(claimant != "James Hauser") %>%
  filter(claimant != "John Patrick") %>%
  filter(claimant != "Kevin Downing") %>%
  filter(claimant != "Lawrence O'Donnell") %>%
  filter(claimant != "Lisa Moore") %>%
  filter(claimant != "Maria Bartiromo") %>%
  filter(claimant != "Mark Levin") %>%
  filter(claimant != "Michael Screnock") %>%
  filter(claimant != "Mike Crute") %>%
  filter(claimant != "Nicholas Burns") %>%
  filter(claimant != "Ray Cross") %>%
  filter(claimant != "Robert Sanborn") %>%
  filter(claimant != "Shelley Grogan")
v7_pf$claimant_party[v7_pf$claimant == "Clay Aiken"] <- "Democratic"
v7_pf$claimant_party[v7_pf$claimant == "Alejandro Mayorkas"] <- "Democratic"
v7_pf$claimant_party[v7_pf$claimant == "Alexander Strenger"] <- "Independent"
v7_pf$claimant_party[v7_pf$claimant == "Beth Parlato"] <- "Republican"
v7_pf$claimant_party[v7_pf$claimant == "Bob Donovan"] <- "Independent"
v7_pf$claimant_party[v7_pf$claimant == "Bob Spindell"] <- "Republican"
v7_pf$claimant_party[v7_pf$claimant == "Brett Giroir"] <- "Republican"
v7_pf$claimant_party[v7_pf$claimant == "Brett McGurk"] <- "Republican"
v7_pf$claimant_party[v7_pf$claimant == "Charles Francis"] <- "Independent"
v7_pf$claimant_party[v7_pf$claimant == "Chris Meagher"] <- "Democratic"
v7_pf$claimant_party[v7_pf$claimant == "George Papadopoulos"] <- "Republican"
v7_pf$claimant_party[v7_pf$claimant == "H.R. McMaster"] <- "Republican"
v7_pf$claimant_party[v7_pf$claimant == "Jeffrey Zients"] <- "Democratic"
v7_pf$claimant_party[v7_pf$claimant == "Jesse Kremer"] <- "Republican"
v7_pf$claimant_party[v7_pf$claimant == "John Kelly"] <- "Republican"
v7_pf$claimant_party[v7_pf$claimant == "John Kirby"] <- "Democratic"
v7_pf$claimant_party[v7_pf$claimant == "Kirstjen Nielsen"] <- "Republican"
v7_pf$claimant_party[v7_pf$claimant == "Loretta Lynch"] <- "Democratic"
v7_pf$claimant_party[v7_pf$claimant == "Louis Marinelli"] <- "Republican"
v7_pf$claimant_party[v7_pf$claimant == "Lowell Holtz"] <- "Independent"
###THIS IS A DUPLICATE CLAIM
v7_pf$claimant_party[v7_pf$claimant == "Marc Butler"] <- "Republican"
v7_pf$claimant_party[v7_pf$claimant == "Michael Long"] <- "Independent"
v7_pf$claimant_party[v7_pf$claimant == "Mike Collier"] <- "Democratic"
v7_pf$claimant_party[v7_pf$claimant == "Nate McMurray"] <- "Democratic"
v7_pf$claimant_party[v7_pf$claimant == "Paul Maner"] <- "Republican"
v7_pf$claimant_party[v7_pf$claimant == "Ryan Frazier"] <- "Republican"
v7_pf$claimant_party[v7_pf$claimant == "Scott Dawson"] <- "Republican"
v7_pf$claimant_party[v7_pf$claimant == "Scott Jones"] <- "Republican"
v7_pf$claimant_party[v7_pf$claimant == "Sidney Powell"] <- "Republican"
v7_pf$claimant_party[v7_pf$claimant == "State representatives"] <- "Republican"
v7_pf$claimant_party[v7_pf$claimant == "Steve Cortes"] <- "Republican"
v7_pf$claimant_party[v7_pf$claimant == "Todd Wilcox"] <- "Republican"
v7_pf$claimant_party[v7_pf$claimant == "William Taylor"] <- "Independent"

v7_pf %>%
  count(claimant_party) %>%
  kable(caption = "V7, PolitiFact, corrected party affiliation")
```

#### The Washington Post unknown affiliation review

```{r wapo unknown affiliation table, message=FALSE, echo= FALSE, warning=FALSE}
v6_wapo %>%
  filter(claimant_party == "unknown_affiliation") %>%
  group_by(claimant) %>% 
  count(sort = TRUE) %>%
  kable(caption = "V6, Washington Post, including unknown affiliation")
```

There were 9 unique claimants and 12 claims.

**The following we assigned a Republican affiliation:**

Charmaine Yoest, [Trump admin appointee](https://en.wikipedia.org/wiki/Charmaine_Yoest), (1 claim)

H.R. McMaster, [National Security Advisor in Trump admin](https://en.wikipedia.org/wiki/H._R._McMaster), (1 claim)

Jim Mattis, [Trump Sec. of Defense](https://en.wikipedia.org/wiki/Jim_Mattis), (1 claim)

Kirstjen Nielsen, [Homeland Security Sec. Trump admin](https://en.wikipedia.org/wiki/Kirstjen_Nielsen), (4 claims)

Steve Cortes, [Former advisor to Donald Trump](https://premierespeakers.com/steve-cortes/bio), (1 claim)

**The following were assigned a Democratic affiliation:**

Vanita Gupta, [worked in the Attorney Gen. office for Obama and Biden](https://en.wikipedia.org/wiki/Vanita_Gupta), (1 claim)

Loretta Lynch, [Attorney General for Obama](https://en.wikipedia.org/wiki/Loretta_Lynch), (1 claim)

**The following were removed for not being political figures:**

Anthony Fauci, [not a political figure](https://en.wikipedia.org/wiki/Anthony_Fauci) (1 claim)

Jim Acosta, [Political commentator](https://en.wikipedia.org/wiki/Jim_Acosta), not a political figure (1 claim)

**In the end, 8 claims were assigned a Republican affiliation, 2 claims were assigned a Democratic affiliation and 2 claims were removed from the Washington Post dataset.** Table 39 shows these updated counts.

```{r v7 substitutions WP, message=FALSE, warning=FALSE, echo=FALSE}
v7_wapo <- v6_wapo %>%
  filter(claimant != "Anthony Fauci") %>%
  filter(claimant != "Jim Acosta")
v7_wapo$claimant_party[v7_wapo$claimant == "Charmaine Yoest"] <- "Republican"
v7_wapo$claimant_party[v7_wapo$claimant == "H.R. McMaster"] <- "Republican"
v7_wapo$claimant_party[v7_wapo$claimant == "Jim Mattis"] <- "Republican"
v7_wapo$claimant_party[v7_wapo$claimant == "Kirstjen Nielsen"] <- "Republican"
v7_wapo$claimant_party[v7_wapo$claimant == "Loretta Lynch"] <- "Democratic"
v7_wapo$claimant_party[v7_wapo$claimant == "Steve Cortes"] <- "Republican"
v7_wapo$claimant_party[v7_wapo$claimant == "Vanita Gupta"] <- "Democratic"

v7_wapo %>%
  count(claimant_party) %>%
  kable(caption = "V7, Washington Post, corrected party affiliation")
```

### Version 8 - Removing Libertarian and Independents

At this stage, it became clear that the Independent and Libertarian claims made up a very small part of each publisher dataset.

These categories were too small to draw statistically relevant conclusions from the data, and so to not distract from potential findings about the two major US parties, claims with an Independent or Libertarian party affiliation were removed from our analysis.
Only Democratic and Republican claims remain in the datasets.

```{r v8, message=FALSE, warning=FALSE, echo=FALSE}
v8_fc <- v7_fc %>%
  filter(claimant_party == "Democratic" |
           claimant_party == "Republican")

v8_fc %>%
  count(claimant_party) %>%
  kable(caption = "V8, FactCheck.Org, Sorted by Claimant Party")

v8_nyt <- v7_nyt %>%
  filter(claimant_party == "Democratic" |
           claimant_party == "Republican")

v8_nyt %>%
  count(claimant_party) %>%
  kable(caption = "V8, New York Times, Sorted by Claimant Party")

v8_pf <- v7_pf %>%
  filter(claimant_party == "Democratic" |
           claimant_party == "Republican")

v8_pf %>%
  count(claimant_party) %>%
  kable(caption = "V8, PolitiFact, Sorted by Claimant Party")

v8_wapo <- v7_wapo %>%
  filter(claimant_party == "Democratic" |
           claimant_party == "Republican")

v8_wapo %>%
  count(claimant_party) %>%
  kable(caption = "V8, Washington Post, Sorted by Claimant Party")
```

At this stage we moved forward with cleaning publisher datasets individually.
Each publisher had it own rating scale, rating quirks, and duplicationss that needed to be addressed.
Also we decided not to move forward with analysis of FactCheck.Org or the New York Times.
To see futher data cleaning or to understand why we did not move forward with our analysis see the following documents in GitHub:\
\
PolitiFact_CleanUp

PolitiFact_Analysis

Washington-Post-Data-Analysis

FactCheck_CleanUp

NewYorkTimes_CleanUp

```{r write v8 csv, message=FALSE, warning=FALSE, echo=FALSE}
write_csv(v8_fc, file = "v8fc.csv")
write_csv(v8_nyt, file = "v8nyt.csv")
write_csv(v8_pf, file = "v8pf.csv")
write_csv(v8_wapo, file = "v8wapo.csv")
```
