---
title: "Location of the Lie"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

```{r loadpackages, echo = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(knitr)
library(kableExtra)
```

```{r echo = FALSE}
loc_tags <- read.csv("pf_mega_location.csv")

loc_tags$textualRating <- factor(loc_tags$textualRating, 
                              levels = c("Pants on Fire", 
                                         "False", 
                                         "Mostly False", 
                                         "Half True", 
                                         "Mostly True", 
                                         "True"))

truenesscolors = c("#FF0000", "#FF872C", "#FAB733", "#FDE64B", "#ACB334", "#69B34C")
```

## Introduction

**Data**: The data used in this analysis is the body of claims fact checked by PolitiFact made by Republican or Democratic political figures between Jan 1st 2016 and June 30th 2021, excluding all statements made by Donald Trump. Further details on how this data was collected and cleaned can be found in the Data Clean Up and PolitiFact Clean Up RMDs and PDFs.

We tagged for two new variables, `location` and `location.extra` by the following definitions. Not all claims were tagged by `location.extra`.

**Advertisement:** campaign ads including video, television, radio and attack ads

**Debate:** campaign debates, panel/roundtable discussion with legislators/tv anchors/etc, back and forth between legislators, debates over legslation.

-   *Location.extra:* [Election Debate]{.ul} or [Unspecified Debate]{.ul}

**Interview:** general interview tag not affiliated with a TV station, radio interview, conversation with reporters, and podcasts that are not their own, general "statement to media" tags by PolitiFact

**Public Event:** town hall, forum, answering questions from the public, roundtable discussions with voters/the public

**Social Media**: posts made on social media.

-   *Location.extra:* platform on which the statement was made ([Facebook]{.ul}, [Instagram]{.ul}, [LinkedIn]{.ul}, [Twitter]{.ul}, [Youtube]{.ul})

**Speech/Remarks:** catchall definition for prepared speeches, remarks, and statements, unless clearly indicated that the speaker is responding to an audience or reporter question; will include a televised address as that is more speech than tv statement; include press briefings/conferences.

-   *Location.extra:* [Campaign Events]{.ul} (which does not necessarily have to be by someone campaigning for themselves, campaigning on the behalf of others will also be tagged as a campaign event), [Hearings]{.ul}, [Press Events]{.ul}, and [Unspecified]{.ul} speeches or remarks.

**TV Appearance:** any televised interview, pre-prepared video remarks.

-   *Location.extra*: the TV channel the claim was made on

**Written Statement:** own website, emails, fliers, own writing/blog posts, letters, mailers, press releases, written policy proposals by candidates.

-   *Location.extra:* [Campaign]{.ul} related material, [Press Releases]{.ul}

## **Methodology**

In the fact check itself, viewed by the URL included in the ClaimReview dataset, PolitiFact will often provide a comment or tag about where the claim was made. For example, a fact check may say: "Amy Klobuchar stated on October 30, 2022 in an interview on Meet the Press..." If the tag given by PolitiFact reasonably falls within one of the aforementioned categories, it was tagged as-is.

If the tag was unclear or additional information is missing, the tagger read the article to determine more details. This is often the case with TV interviews, with general "statements" etc.

Only in extremely unclear circumstances taggers went beyond the article, which usually involved looking up the TV channel on which a particular show is broadcasted, or using PolitiFact's citations to find the original location of the lie. If tagged as something in the header and something else in the text, taggers were biased towards the original tag that PolitiFact gave, unless the article presents significant evidence that the location of the lie is not explained accurately by the header (for our purposes).

In an effort to reduce the number of judgment calls made, we recorded potential areas of confusion or gray areas to explain tagging logic. These are noted below.

**Advertisement:** Campaign videos are tagged as advertisements, as this is the category that best encompasses their nature.

**Debate:** The `location.extra` tag [Election Debate]{.ul} refers to any debate between competing candidates. If the PolitiFact url indicates the location as a debate, but does not specify the type/presence of an opposing candidate, then it is not tagged as [Election Debate]{.ul}. We recognize this may mean that some data is omitted from our query.

**Interview:** Facebook live interviews get tagged as interviews over social media, to reflect the type of speech in which the statement was made---having the purpose of determining the context of the lie, the Interview tag is more representative than the Social Media tag. This is the same for interviews published as YouTube videos.

**Social Media:** Video commentary on a certain issue posted by a politician on Facebook was tagged as a Social Media lie because the candidate created, reviewed, and decided to post it on social media. We found this to be intentionally different from the definition of Speech/Remarks, and TV Appearance, which were the other potential tags for this claim.

**Speech/Remarks:** The `location.extra` tag, [Campaign,]{.ul} is inherently complex due to the methodology of tagging. We found that it became difficult to tag for all comments made by a candidate during a campaign, regardless of whether it is an official campaign or not. PolitiFact does not consistently identify an event as a campaign event, and there are situations in which candidates for office hold spontaneous and thus unofficial public events, not prepared as official campaign events. Taggers made every attempt to tag consistently, but there is a certain amount of unavoidable subjectivity to the PolitiFact tags; the inconsistent amount of information that they provide may have been compounded by our own biases and subjectivity. We set tagging guidelines to mitigate the highest number of judgement calls possible. All speeches or remarks made at the RNC and DNC are tagged as [Campaign]{.ul} regardless of whether the speaker is a candidate or not. Since we were tagging for the location of the lie and not the speaker, this is a campaign event and the tag reflects that---claims made by individuals campaigning on behalf of someone else or during a campaign event were still given this tag. If a person had not yet announced their candidature at the time of the claim, they were not tagged as campaign statements. If a speech was not explicitly listed as a campaign event but the fact-check began with "person X, candidate for/running for..." then it was tagged as campaign unless there was a specified alternative purpose for the event.

Also, there were certain instances where Speech/Remarks applied to a claim along with several other tags. In each of these instances, we tried to tag for the nature of the speech. *For example*, despite being spoken on the radio, Rob Swearingen's weekly radio address was tagged as Speech/Remarks over Interview because a radio address has more components of a speech (preparation, monologue, etc) than a radio interview (off-the-cuff, dynamic, conversation between two individuals, etc), despite radio interview to better reflect the type of speech within which the lie was contained.

**TV Appearance:** If a televised event is also something else (a town hall, a debate, a speech, etc), the other aspect of the speech takes precedence. *For example,* a televised town hall will be tagged as a town hall, because the context in which the lie was said was the town tall, the fact that it was televised does not change that situation. On the contrary, if a TV channel hosts a casual round-table discussion that is exclusively displayed on that channel, that will be tagged as a TV Appearance on that channel.

Also, in cases where it was not explicitly stated whether an interview was conducted on TV or not, we used contextual clues to make determinations, reading the article for further information. The most common clue was when the interviewer was a major news anchor or TV personality. In these cases, the claim was tagged as a TV appearance.

In `location.extra` the Regional Channel tag became slightly complex, unaffiliated local TV stations are tagged as Regional Channel, whereas local programs on national news networks are tagged by the parent organization. *For example,* WISN-TV, a channel that only plays in Milwaukee, Wisconsin, will be tagged as Regional Channel whereas CBS Miami's "Facing South Florida" will be tagged as CBS.

**Written Statement:** The [Campaign]{.ul} tag under `location.extra` also requires further specification. Policy proposals written by candidates in advance of an election were be tagged as campaign material, as were endorsements and websites made by a candidate to attack their opponent in an election.

Fact sheets are tagged in `location.extra` as Press Releases, as we find their intent, production, and distribution to be sufficiently similar to equate them.

Things that were reported as having been repeated via various platforms, where at least some of the recorded versions of the lie are in written form are tagged as "Written Statement" under `location` and then "Various Statements" in `location.extra` because that represents the highest level of vetting the statement may have gone through and therefore the greatest degree of intentionality.

Talking points written by the campaign and given to staff to be read on behalf of the candidate's campaign are tagged as Written Statement, as that similarly represents the high degree of oversight these statements could have received.

\pagebreak

## Data Distribution by Party

```{r, echo = FALSE}
loc_tags %>%
  group_by(claimant_party) %>%
  summarise(count = n(), percentage = (100*count)/nrow(loc_tags)) %>%
  kable(caption = "Claim Counts by Party", 
        digits = 2,
        format = "pipe")
```

In this dataset, there were approximately 100 more claims made by Democrats versus Republicans. Nevertheless, given the overall claim counts, the ratio of Democrat's claims to Republican's is $1.09:1$, allowing us to compare the two.

```{r, echo = FALSE, fig.height = 4}
loc_tags %>%
  ggplot(aes(x = claimant_party, fill = textualRating)) +
  geom_bar(position = "fill") +
  labs(
    x = "Claimant Party",
    y = "Rating Distribution",
    fill = "PolitiFact Rating",
    title = "Claim Ratings by Political Party"
  ) +
  scale_fill_manual(values = truenesscolors)
```

\pagebreak

## General Analysis of Location Tag

```{r, echo = FALSE}
loc_tags %>%
  group_by(location) %>%
  summarise(count = n(), percentage = (n()*100)/nrow(loc_tags)) %>%
  arrange(desc(count)) %>%
  kable(caption = "Distribution of Claims by Location", 
        digits = 1,
        format = "pipe")
```

The three largest categories of claims in the PolitiFact database are Speech/Remarks, TV Appearance, and Social Media. The smallest categories were Public Event, Advertisement, and Interview.

```{r, echo = FALSE, fig.height = 3}
loc_relevel <- loc_tags %>%
  mutate(location = fct_relevel(location, c("Speech/Remarks",
                                            "Debate",
                                            "Written Statement",
                                            "Social Media",
                                            "Public Event",
                                            "TV Appearance",
                                            "Interview",
                                            "Advertisement")))

loc_relevel %>%
  ggplot(aes(x = location, fill = textualRating)) +
  geom_bar() +
  labs(
    x = "Location",
    y = "Rating Distribution",
    fill = "PolitiFact Rating",
    title = "Claim Ratings, Count by Location"
  ) +
  scale_fill_manual(values = truenesscolors) +
  theme(axis.text.x = element_text(size = 5))  

loc_relevel %>%
  ggplot(aes(x = location, fill = textualRating)) +
  geom_bar(position = "fill") +
  labs(
    x = "Location",
    y = "Rating Distribution",
    fill = "PolitiFact Rating",
    title = "Claim Ratings, Proportionally by Location"
  ) +
  scale_fill_manual(values = truenesscolors) +
  theme(axis.text.x = element_text(size = 5)) 
```

```{r, echo = FALSE}
all_loc_count <- loc_tags %>%
  group_by(location) %>%
  summarize(count = n()) %>%
  pull(count)

all_ad <- loc_tags %>%
  filter(location == "Advertisement") %>%
  group_by(textualRating) %>%
  summarize(percentage = (n()*100)/all_loc_count[1])
all_ad$location <- "Advertisement"

all_deb <- loc_tags %>%
  filter(location == "Debate") %>%
  group_by(textualRating) %>%
  summarize(percentage = (n()*100)/all_loc_count[2])
all_deb$location <- "Debate"

all_int <- loc_tags %>%
  filter(location == "Interview") %>%
  group_by(textualRating) %>%
  summarize(percentage = (n()*100)/all_loc_count[3])
all_int$location <- "Interview"

all_pub <- loc_tags %>%
  filter(location == "Public Event") %>%
  group_by(textualRating) %>%
  summarize(percentage = (n()*100)/all_loc_count[4])
all_pub$location <- "Public Event"

all_soc <- loc_tags %>%
  filter(location == "Social Media") %>%
  group_by(textualRating) %>%
  summarize(percentage = (n()*100)/all_loc_count[5])
all_soc$location <- "Social Media"

all_spe <- loc_tags %>%
  filter(location == "Speech/Remarks") %>%
  group_by(textualRating) %>%
  summarize(percentage = (n()*100)/all_loc_count[6])
all_spe$location <- "Speech/Remarks"

all_tv <- loc_tags %>%
  filter(location == "TV Appearance") %>%
  group_by(textualRating) %>%
  summarize(percentage = (n()*100)/all_loc_count[7])
all_tv$location <- "TV Appearance"

all_wri <- loc_tags %>%
  filter(location == "Written Statement") %>%
  group_by(textualRating) %>%
  summarize(percentage = (n()*100)/all_loc_count[8])
all_wri$location <- "Written Statement"

all_loc <- rbind(all_ad, all_deb, all_int, all_pub, all_soc, all_spe, all_tv, all_wri)
all_loc %>%
  pivot_wider(names_from = textualRating, values_from = percentage) %>%
  kable(caption = "Percent of Rating Distribution, by Location", 
        digits = 1, 
        format = "pipe")
```

The highest proportion of True-ish (True, Mostly True, Half True) claims are made in Speech/Remarks (62.4%). The lowest proportion of True-ish claims are made in Advertisements (44.2%). The highest proportion of Pants on Fire claims appear in Interviews (8.7%). The highest proportion of True statements are made at Public Events (18.8%).

```{r, echo = FALSE, message = FALSE}
loc_tags %>%
  group_by(location, claimant_party) %>%
  summarise(count = n()) %>%
  pivot_wider(names_from = location, values_from = count) %>%
  kable(caption = "Distribution of Claims by Party and Location") %>%
  kable_styling(full_width = TRUE, font_size = 7)
```

Comparing the ratio of claim counts by Party and Location, Interview, Advertisement, and Social Media have nearly a $1:1$ ratio of Democratic to Republican claims. Public Event has the most egregious difference, with a ratio of $2.48:1$ Democratic to Republican claims. Debate, Speech/Remarks, TV Appearance, and Written Statements have no greater than a $1.5:1$ ratio.

Some of these differences, namely the $1.5:1$ ratio of Democratic to Republican claims in the Debate category may be attributable to the omission of Trump from the dataset, as all of Clinton's claims from their 2016 debates are included, but his are not.

```{r, echo = FALSE, fig.height = 9}
loc_relevel %>%
  ggplot(aes(x = claimant_party, fill = textualRating)) +
  geom_bar(position = "fill") +
  facet_wrap(~ location, ncol = 2) +
  labs(
    x = "Claimant Party",
    y = "Rating Distribution",
    fill = "PolitiFact Rating",
    title = "Claim Ratings by Political Party and Location"
  ) +
  scale_fill_manual(values = truenesscolors)
```

\newpage

## **Social Media**

```{r, echo = FALSE}
loc_tags %>%
  filter(location == "Social Media") %>%
  group_by(location.extra) %>%
  count() %>%
  arrange(desc(n)) %>%
  kable(caption = "Social Media Tag, Distribution of Claims Platform",
        format = "pipe")
```

Seeing that the vast majority of claims originate on Twitter and Facebook, we focused on those. The other platforms did not have sufficient claims to justify analysis.

```{r, echo = FALSE, fig.height = 4}
loc_tags %>%
  filter(location == "Social Media",
         location.extra == "Twitter" | location.extra == "Facebook") %>%
  ggplot(aes(x = location.extra, fill = textualRating)) +
  geom_bar(position = "fill") +
  labs(
    x = "Social Media Platform",
    y = "Rating Distribution",
    fill = "PolitiFact Rating",
    title = "Claim Ratings by Platform"
  ) +
  scale_fill_manual(values = truenesscolors)
```

```{r, echo = FALSE}
social <- loc_tags %>%
  filter(location == "Social Media",
         location.extra == "Twitter" | location.extra == "Facebook") %>%
  group_by(location.extra) %>%
  summarize(count = n()) %>%
  pull(count)

social_fac <- loc_tags %>%
  filter(location == "Social Media",
         location.extra == "Facebook") %>%
  group_by(textualRating) %>%
  summarize(percentage = (100*n())/social[1]) %>%
  mutate(Total = sum(percentage))
social_fac$platform <- "Facebook"

social_twit <- loc_tags %>%
  filter(location == "Social Media",
         location.extra == "Twitter") %>%
  group_by(textualRating) %>%
  summarize(percentage = (100*n())/social[2]) %>%
  mutate(Total = sum(percentage))
social_twit$platform <- "Twitter"

social_prop <- rbind(social_fac, social_twit) 

social_prop %>%
  pivot_wider(names_from = textualRating, values_from = percentage) %>%
  kable(caption = "Percentage of Ratings by Platform", 
        digits = 1,
        format = "pipe")
```

This data appears to indicate that Facebook has a higher proportion of falsehoods than Twitter, among all checked statements on PolitiFact. Of all statements on Facebook, 50.1% of statements were found to be Mostly False, False, or Pants on Fire, compared to 39.7% of the same on Twitter. Twitter also had a higher proportion of True statements checked, 16.4% to Facebook's 12.9%, about 1.27x higher. There are proportionally 2.08x as many Pants on Fire claims on Facebook when compared to Twitter.

We took particular interest in the party distribution on Twitter, a platform often regarded as the primary online public forum for discussion.

```{r, echo = FALSE, fig.height = 4}
loc_tags %>%
  filter(location == "Social Media",
         location.extra == "Twitter") %>%
  ggplot(aes(x = claimant_party, fill = textualRating)) +
  geom_bar(position = "fill") +
  labs(
    x = "Political Party",
    y = "Rating Distribution",
    fill = "PolitiFact Rating",
    title = "Claim Ratings on Twitter by Party"
  ) +
  scale_fill_manual(values = truenesscolors)

loc_tags %>%
  filter(location == "Social Media",
         location.extra == "Twitter") %>%
  group_by(claimant_party) %>%
  summarise(count = n()) %>%
  pivot_wider(names_from = claimant_party, values_from = count) %>%
  kable(caption = "Twitter Claims by Party",
        format = "pipe")
```

```{r, echo = FALSE}
twit_claims <- loc_tags %>%
  filter(location == "Social Media",
         location.extra == "Twitter") %>%
  group_by(claimant_party) %>%
  summarise(count = n()) %>%
  pull(count)

twit_dems <- loc_tags %>%
  filter(location == "Social Media",
         location.extra == "Twitter",
         claimant_party == "Democratic") %>%
  group_by(textualRating) %>%
  summarise(percentage = (100*n())/twit_claims[1])
twit_dems$party <- "Democratic"

twit_reps <- loc_tags %>%
  filter(location == "Social Media",
         location.extra == "Twitter",
         claimant_party == "Republican") %>%
  group_by(textualRating) %>%
  summarise(percentage = (100*n())/twit_claims[2])
twit_reps$party <- "Republican"

twit_props <- rbind(twit_dems, twit_reps)
twit_props %>%
  pivot_wider(names_from = textualRating, values_from = percentage) %>%
  kable(caption = "Percentage of Twitter Claims by Rating and Party",
        digits = 1,
        format = "pipe")

twit_dems_count <- loc_tags %>%
  filter(location == "Social Media",
         location.extra == "Twitter",
         claimant_party == "Democratic") %>%
  group_by(textualRating) %>%
  summarise(count = n())
twit_dems_count$party <- "Democratic"

twit_reps_count <- loc_tags %>%
  filter(location == "Social Media",
         location.extra == "Twitter",
         claimant_party == "Republican") %>%
  group_by(textualRating) %>%
  summarise(count = n())
twit_reps_count$party <- "Republican"

twit_count <- rbind(twit_dems_count, twit_reps_count)
twit_count %>%
  pivot_wider(names_from = textualRating, values_from = count, values_fill = 0) %>%
  kable(caption = "Count of Twitter Claims by Rating and Party",
        digits = 1,
        format = "pipe")
```

\newpage

## **TV Appearance**

```{r, echo = FALSE}
loc_tags %>%
  filter(location == "TV Appearance") %>%
  group_by(location.extra) %>%
  count() %>%
  arrange(desc(n)) %>%
  kable(caption = "Distribution of TV Appearances",
        format = "pipe")
```

For the analysis on this tag, we focused on all TV channels on which at least 30 claims were made. This leaves: ABC, CBS, CNN, FOX, MSNBC, NBC, and the Regional Station tag.

```{r, echo = FALSE, fig.height = 4}
tv_cleaned <- loc_tags %>%
  filter(location == "TV Appearance",
         location.extra == "FOX" | 
           location.extra == "ABC" |
           location.extra == "CBS" |
           location.extra == "CNN" | 
           location.extra == "MSNBC" | 
           location.extra == "NBC" | 
           location.extra == "Regional Station")

tv_relevel <- tv_cleaned %>%
  mutate(location.extra = fct_relevel(location.extra, c("FOX",
                                                        "NBC",
                                                        "Regional Station",
                                                        "CNN",
                                                        "MSNBC",
                                                        "ABC",
                                                        "CBS")))
tv_relevel %>%
  ggplot(aes(x = location.extra, fill = textualRating)) +
  geom_bar(position = "fill") +
  labs(
    x = "TV Channel",
    y = "Rating Distribution",
    fill = "PolitiFact Rating",
    title = "Claim Ratings by TV Channel"
  ) +
  scale_fill_manual(values = truenesscolors)
```

```{r, echo = FALSE, message = FALSE}
tv_totals <- tv_cleaned %>%
  group_by(location.extra) %>%
  summarize(count = n()) %>%
  pull(count)

abc_prop <- loc_tags %>%
  filter(location == "TV Appearance",
         location.extra == "ABC") %>%
  group_by(textualRating) %>%
  summarize(percentage = (100*n())/tv_totals[1])
abc_prop$channel <- "ABC"

cbs_prop <- loc_tags %>%
  filter(location == "TV Appearance",
         location.extra == "CBS") %>%
  group_by(textualRating) %>%
  summarize(percentage = (100*n())/tv_totals[2])
cbs_prop$channel <- "CBS"

cnn_prop <- loc_tags %>%
  filter(location == "TV Appearance",
         location.extra == "CNN") %>%
  group_by(textualRating) %>%
  summarize(percentage = (100*n())/tv_totals[3])
cnn_prop$channel <- "CNN"

fox_prop <- loc_tags %>%
  filter(location == "TV Appearance",
         location.extra == "FOX") %>%
  group_by(textualRating) %>%
  summarize(percentage = (100*n())/tv_totals[4])
fox_prop$channel <- "FOX"

msnbc_prop <- loc_tags %>%
  filter(location == "TV Appearance",
         location.extra == "MSNBC") %>%
  group_by(textualRating) %>%
  summarize(percentage = (100*n())/tv_totals[5])
msnbc_prop$channel <- "MSNBC"

nbc_prop <- loc_tags %>%
  filter(location == "TV Appearance",
         location.extra == "NBC") %>%
  group_by(textualRating) %>%
  summarize(percentage = (100*n())/tv_totals[6])
nbc_prop$channel <- "NBC"
  
rs_prop <- loc_tags %>%
  filter(location == "TV Appearance",
         location.extra == "Regional Station") %>%
  group_by(textualRating) %>%
  summarize(percentage = (100*n())/tv_totals[7])
rs_prop$channel <- "Regional Station"

tv_props <- rbind(abc_prop, cbs_prop, cnn_prop, fox_prop, msnbc_prop, nbc_prop, rs_prop)

tv_props %>%
  pivot_wider(names_from = textualRating, values_from = percentage) %>%
  kable(caption = "Claim Rating Percentages by TV Channel", 
        digits = 1,
        format = "pipe")
```

Basing trueness on Half True, Mostly True, and True statements, the descending order of trueness by TV Channel goes as follows: CBS (62.8%), MSNBC (61.6% True-ish), CNN (57.4% True-ish), Regional Stations (56.2% True-ish), NBC (45.2% True-ish), ABC (42.6% True-ish and finally FOX (38.9% True-ish).

Among these news channels, CNN, MSNBC, and NBC had nearly identical proportions of Pants on Fire claims, 5.9%, 5.0%, and 5.5% respectively.

```{r, echo = FALSE, fig.height = 8}
tv_relevel %>%
  ggplot(aes(x = claimant_party, fill = textualRating)) +
  geom_bar(position = "fill") +
  facet_wrap(~ location.extra, ncol = 2) +
  labs(
    x = "Claimant Party",
    y = "Rating Distribution",
    fill = "PolitiFact Rating",
    title = "Claim Ratings by Political Party and TV Channel"
  ) +
  scale_fill_manual(values = truenesscolors)
```

\newpage

```{r, echo = FALSE, message = FALSE}
abc_count <- loc_tags %>%
   filter(location == "TV Appearance",
         location.extra == "ABC") %>%
  group_by(claimant_party, textualRating) %>%
  summarize(count = n())
abc_count$channel <- "ABC"

cbs_count <- loc_tags %>%
   filter(location == "TV Appearance",
         location.extra == "CBS") %>%
  group_by(claimant_party, textualRating) %>%
  summarize(count = n()) 
cbs_count$channel <- "CBS"

cnn_count <- loc_tags %>%
   filter(location == "TV Appearance",
         location.extra == "CNN") %>%
  group_by(claimant_party, textualRating) %>%
  summarize(count = n())
cnn_count$channel <- "CNN"

fox_count <- loc_tags %>%
   filter(location == "TV Appearance",
         location.extra == "FOX") %>%
  group_by(claimant_party, textualRating) %>%
  summarize(count = n())
fox_count$channel <- "FOX"

msnbc_count <- loc_tags %>%
   filter(location == "TV Appearance",
         location.extra == "MSNBC") %>%
  group_by(claimant_party, textualRating) %>%
  summarize(count = n())
msnbc_count$channel <- "MSNBC"

nbc_count <- loc_tags %>%
   filter(location == "TV Appearance",
         location.extra == "NBC") %>%
  group_by(claimant_party, textualRating) %>%
  summarize(count = n())
nbc_count$channel <- "NBC"

reg_count <- loc_tags %>%
   filter(location == "TV Appearance",
         location.extra == "Regional Station") %>%
  group_by(claimant_party, textualRating) %>%
  summarize(count = n()) 
reg_count$channel <- "Regional Station"

count <- rbind(abc_count, 
               cbs_count, 
               cnn_count, 
               fox_count, 
               msnbc_count, 
               nbc_count, 
               reg_count) %>%
  mutate(party = claimant_party) %>%
  ungroup() %>%
  select(-1)

count %>%
  pivot_wider(names_from = textualRating, values_from = count, values_fill = 0) %>%
  kable(caption = "Claim Count by Party and TV Channel",
        format = "pipe")
```

```{r, echo = FALSE, message = FALSE}
tv_totals_dems <- tv_cleaned %>%
  filter(claimant_party == "Democratic") %>%
  group_by(location.extra) %>%
  summarize(count = n()) %>%
  pull(count)

abc_prop_dems <- loc_tags %>%
  filter(location == "TV Appearance",
         location.extra == "ABC",
         claimant_party == "Democratic") %>%
  group_by(textualRating) %>%
  summarize(percentage = (100*n())/tv_totals_dems[1])
abc_prop_dems$channel <- "ABC"
abc_prop_dems$party <- "Democratic"

cbs_prop_dems <- loc_tags %>%
  filter(location == "TV Appearance",
         location.extra == "CBS",
         claimant_party == "Democratic") %>%
  group_by(textualRating) %>%
  summarize(percentage = (100*n())/tv_totals_dems[2])
cbs_prop_dems$channel <- "CBS"
cbs_prop_dems$party <- "Democratic"

cnn_prop_dems <- loc_tags %>%
  filter(location == "TV Appearance",
         location.extra == "CNN",
         claimant_party == "Democratic") %>%
  group_by(textualRating) %>%
  summarize(percentage = (100*n())/tv_totals_dems[3])
cnn_prop_dems$channel <- "CNN"
cnn_prop_dems$party <- "Democratic"

fox_prop_dems <- loc_tags %>%
  filter(location == "TV Appearance",
         location.extra == "FOX",
         claimant_party == "Democratic") %>%
  group_by(textualRating) %>%
  summarize(percentage = (100*n())/tv_totals_dems[4])
fox_prop_dems$channel <- "FOX"
fox_prop_dems$party <- "Democratic"

msnbc_prop_dems <- loc_tags %>%
  filter(location == "TV Appearance",
         location.extra == "MSNBC",
         claimant_party == "Democratic") %>%
  group_by(textualRating) %>%
  summarize(percentage = (100*n())/tv_totals_dems[5])
msnbc_prop_dems$channel <- "MSNBC"
msnbc_prop_dems$party <- "Democratic"

nbc_prop_dems <- loc_tags %>%
  filter(location == "TV Appearance",
         location.extra == "NBC",
         claimant_party == "Democratic") %>%
  group_by(textualRating) %>%
  summarize(percentage = (100*n())/tv_totals_dems[6])
nbc_prop_dems$channel <- "NBC"
nbc_prop_dems$party <- "Democratic"
  
rs_prop_dems <- loc_tags %>%
  filter(location == "TV Appearance",
         location.extra == "Regional Station",
         claimant_party == "Democratic") %>%
  group_by(textualRating) %>%
  summarize(percentage = (100*n())/tv_totals_dems[7])
rs_prop_dems$channel <- "Regional Station"
rs_prop_dems$party <- "Democratic"

tv_props_dems <- rbind(abc_prop_dems, cbs_prop_dems, cnn_prop_dems, fox_prop_dems, msnbc_prop_dems, nbc_prop_dems, rs_prop_dems)

tv_totals_reps <- tv_cleaned %>%
  filter(claimant_party == "Republican") %>%
  group_by(location.extra) %>%
  summarize(count = n()) %>%
  pull(count)

abc_prop_reps <- loc_tags %>%
  filter(location == "TV Appearance",
         location.extra == "ABC",
         claimant_party == "Republican") %>%
  group_by(textualRating) %>%
  summarize(percentage = (100*n())/tv_totals_reps[1])
abc_prop_reps$channel <- "ABC"
abc_prop_reps$party <- "Republican"

cbs_prop_reps <- loc_tags %>%
  filter(location == "TV Appearance",
         location.extra == "CBS",
         claimant_party == "Republican") %>%
  group_by(textualRating) %>%
  summarize(percentage = (100*n())/tv_totals_reps[2])
cbs_prop_reps$channel <- "CBS"
cbs_prop_reps$party <- "Republican"

cnn_prop_reps <- loc_tags %>%
  filter(location == "TV Appearance",
         location.extra == "CNN",
         claimant_party == "Republican") %>%
  group_by(textualRating) %>%
  summarize(percentage = (100*n())/tv_totals_reps[3])
cnn_prop_reps$channel <- "CNN"
cnn_prop_reps$party <- "Republican"

fox_prop_reps <- loc_tags %>%
  filter(location == "TV Appearance",
         location.extra == "FOX",
         claimant_party == "Republican") %>%
  group_by(textualRating) %>%
  summarize(percentage = (100*n())/tv_totals_reps[4])
fox_prop_reps$channel <- "FOX"
fox_prop_reps$party <- "Republican"

msnbc_prop_reps <- loc_tags %>%
  filter(location == "TV Appearance",
         location.extra == "MSNBC",
         claimant_party == "Republican") %>%
  group_by(textualRating) %>%
  summarize(percentage = (100*n())/tv_totals_reps[5])
msnbc_prop_reps$channel <- "MSNBC"
msnbc_prop_reps$party <- "Republican"

nbc_prop_reps <- loc_tags %>%
  filter(location == "TV Appearance",
         location.extra == "NBC",
         claimant_party == "Republican") %>%
  group_by(textualRating) %>%
  summarize(percentage = (100*n())/tv_totals_reps[6])
nbc_prop_reps$channel <- "NBC"
nbc_prop_reps$party <- "Republican"
  
rs_prop_reps <- loc_tags %>%
  filter(location == "TV Appearance",
         location.extra == "Regional Station",
         claimant_party == "Republican") %>%
  group_by(textualRating) %>%
  summarize(percentage = (100*n())/tv_totals_reps[7])
rs_prop_reps$channel <- "Regional Station"
rs_prop_reps$party <- "Republican"

tv_props_reps <- rbind(abc_prop_reps, cbs_prop_reps, cnn_prop_reps, fox_prop_reps, msnbc_prop_reps, nbc_prop_reps, rs_prop_reps)

tv_props_byparty <- rbind(tv_props_dems, tv_props_reps)

tv_props_byparty %>%
  arrange(channel) %>%
  pivot_wider(names_from = textualRating, values_from = percentage, values_fill = 0) %>%
  kable(caption = "Claim Rating Proportions by Party and TV Channel", 
        digits = 1,
        format = "pipe")
```
